---
title: An R Markdown document converted from "AC2/r-debiased-ml-for-partially-linear-iv-model.irnb"
output: html_document
---

# Double/Debiased ML for Partially Linear IV Model

This is a simple implementation of Debiased Machine Learning for the Partially Linear
IV Regression Model, which provides an application of DML IV inference.


Reference:

- https://arxiv.org/abs/1608.00060
- https://www.amazon.com/Business-Data-Science-Combining-Accelerate/dp/1260452778

The code is based on the book.


# Partially Linear IV Model

We consider the partially linear structural equation model:
\begin{align}
 Y :=~& D\theta_0 + g_0(X) + \zeta,  & E[\zeta \mid Z,X]= 0,\\
 Z :=~& m_0(X) +  V,   &  E[V \mid X] = 0.
\end{align}


Note that this model is not a regression model unless $Z=D$.  The model  is a canonical
model in causal inference, going back to P. Wright's work on IV methods for estimating demand/supply equations, with the modern difference being that $g_0$ and $m_0$ are nonlinear, potentially complicated functions of high-dimensional $X$.  


The idea of this model is that there is a structural or causal relation between $Y$ and $D$, captured by $\theta_0$, and $g_0(X) + \zeta$ is the stochastic error, partly explained by covariates $X$.  $V$ and $\zeta$ are stochastic errors that are not explained by $X$. Since $Y$ and $D$ are jointly determined, we need an external factor, commonly referred to as an instrument, $Z$ to create exogenous variation
in $D$.   Note that $Z$ should affect $D$.  The $X$ here serve again as confounding factors, so we can think of variation in $Z$ as being exogenous only conditional on $X$.


The causal DAG this model corresponds to is given by:
$$
Z \to D,  X \to (Y, Z, D),  L \to (Y,D),
$$
where $L$ is the latent confounder affecting both $Y$ and $D$, but not $Z$.



---

# Example

A simple contextual example is from biostatistics, where $Y$ is a health outcome and $D$ is indicator of smoking.  Thus, $\theta_0$ is captures the effect of smoking on health.  Health outcome $Y$ and smoking behavior $D$ are treated as being jointly determined.  $X$ represents patient characteristics, and $Z$ could be a doctor's advice not to smoke (or another behavioral treatment) that may affect the outcome $Y$ only through shifting the behavior $D$, conditional on characteristics $X$.   

----



# PLIVM in Residualized Form

The PLIV model above can be rewritten in the following residualized form:
$$
  \tilde Y = \tilde D \theta_0 + \zeta,   \quad  E[\zeta \mid V,X]= 0,
$$
where
$$
 \tilde Y = (Y- \ell_0(X)),  \quad \ell_0(X) = E[Y \mid X] \\
   \tilde D = (D - r_0(X)), \quad r_0(X) = E[D \mid X] \\
   \tilde Z = (Z- m_0(X)), \quad m_0(X) = E[Z \mid X].
$$
   The "tilde" variables (e.g. $\tilde Y$) above represent original variables after taking out or "partialling out"
  the effect of $X$.  Note that $\theta_0$ is identified from this equation if $V$
  and $U$ have non-zero correlation, which automatically means that $U$ and $V$
  must have non-zero variation.

  

-----

# DML for PLIV Model

Given identification, DML  proceeds as follows

Compute the estimates $\hat \ell_0$, $\hat r_0$, and $\hat m_0$ , which amounts
to solving the three problems of predicting $Y$, $D$, and $Z$ using
$X$, using any generic  ML method, giving us estimated residuals
$$
\tilde Y = Y - \hat \ell_0(X), \\ \tilde D= D - \hat r_0(X), \\ \tilde Z = Z- \hat m_0(X).
$$
The estimates should be of a cross-validated form, as detailed in the algorithm below.

Estimate $\theta_0$ by the the intstrumental
variable regression of $\tilde Y$ on $\tilde D$ using $\tilde Z$ as an instrument.
Use the conventional inference for the IV regression estimator, ignoring
the estimation error in these residuals.

The reason we work with this residualized form is that it eliminates the bias
arising when solving the prediction problem in stage 1. The role of cross-validation
is to avoid another source of bias due to potential overfitting.

The estimator is adaptive,
in the sense that the first stage estimation errors do not affect the second
stage errors.

```{r}
install.packages("hdm")
install.packages("AER")
install.packages("randomForest")
install.packages("lfe")
install.packages("glmnet")
```

```{r}
library(hdm)
library(AER) # applied econometrics library
library(randomForest) # random Forest library
library(lfe) # high-dimensional econometrics library
library(glmnet) # glm net

set.seed(1)
```

```{r}
# DML for PLIVM

dml2_for_plivm <- function(x, d, z, y, dreg, yreg, zreg, nfold = 5) {
  # this implements DML2 algorithm, where there moments are estimated via DML, before constructing
  # the pooled estimate of theta randomly split data into folds
  nobs <- nrow(x)
  foldid <- rep.int(1:nfold, times = ceiling(nobs / nfold))[sample.int(nobs)]
  I <- split(1:nobs, foldid)
  # create residualized objects to fill
  ytil <- dtil <- ztil <- rep(NA, nobs)
  # obtain cross-fitted residuals
  cat("fold: ")
  for (b in seq_along(I)) {
    dfit <- dreg(x[-I[[b]], ], d[-I[[b]]]) # take a fold out
    zfit <- zreg(x[-I[[b]], ], z[-I[[b]]]) # take a fold out
    yfit <- yreg(x[-I[[b]], ], y[-I[[b]]]) # take a folot out
    dhat <- predict(dfit, x[I[[b]], ], type = "response") # predict the fold out
    zhat <- predict(zfit, x[I[[b]], ], type = "response") # predict the fold out
    yhat <- predict(yfit, x[I[[b]], ], type = "response") # predict the fold out
    dtil[I[[b]]] <- (d[I[[b]]] - dhat) # record residual
    ztil[I[[b]]] <- (z[I[[b]]] - zhat) # record residual
    ytil[I[[b]]] <- (y[I[[b]]] - yhat) # record residial
    cat(b, " ")
  }
  ivfit <- tsls(y = ytil, d = dtil, x = NULL, z = ztil, intercept = FALSE)
  coef_est <- ivfit$coef # extract coefficient
  se <- ivfit$se # record standard error
  cat(sprintf("\ncoef (se) = %g (%g)\n", coef_est, se))
  return(list(coef_est = coef_est, se = se, dtil = dtil, ytil = ytil, ztil = ztil))
}
```

-----

# Emprical Example:  Acemoglu, Johnson, Robinson (AER).


* Y is log GDP;
* D is a measure of Protection from Expropriation, a proxy for
quality of insitutions;
* Z is the log of Settler's mortality;
* W are geographical variables (latitude, latitude squared, continent dummies as well as interactions)


```{r}
file <- "https://raw.githubusercontent.com/CausalAIBook/MetricsMLNotebooks/main/data/AJR.csv"
AJR <- read.csv(file)
dim(AJR)
```

```{r}
y <- AJR$GDP
d <- AJR$Exprop
z <- AJR$logMort
xraw <- model.matrix(~ Latitude + Africa + Asia + Namer + Samer, data = AJR)
x <- model.matrix(~ -1 + (Latitude + Latitude2 + Africa + Asia + Namer + Samer)^2, data = AJR)
dim(x)
```

```{r}
set.seed(1)
# DML with PostLasso
cat(sprintf("\n DML with Post-Lasso \n"))

dreg <- function(x, d) {
  hdm::rlasso(x, d)
} # ML method=lasso
yreg <- function(x, y) {
  hdm::rlasso(x, y)
} # ML method=lasso
zreg <- function(x, z) {
  hdm::rlasso(x, z)
} # ML method=lasso

dml2_lasso <- dml2_for_plivm(x, d, z, y, dreg, yreg, zreg, nfold = 5)


# DML with Random Forest
cat(sprintf("\n DML with Random Forest \n"))

dreg <- function(x, d) {
  randomForest(x, d)
} # ML method=Forest
yreg <- function(x, y) {
  randomForest(x, y)
} # ML method=Forest
zreg <- function(x, z) {
  randomForest(x, z)
} # ML method=Forest

dml2_rf <- dml2_for_plivm(xraw, d, z, y, dreg, yreg, zreg, nfold = 5)

# Compare Forest vs Lasso
comp_tab <- matrix(NA, 3, 2)
comp_tab[1, ] <- c(sqrt(mean((dml2_rf$ytil)^2)), sqrt(mean((dml2_lasso$ytil)^2)))
comp_tab[2, ] <- c(sqrt(mean((dml2_rf$dtil)^2)), sqrt(mean((dml2_lasso$dtil)^2)))
comp_tab[3, ] <- c(sqrt(mean((dml2_rf$ztil)^2)), sqrt(mean((dml2_lasso$ztil)^2)))
rownames(comp_tab) <- c("RMSE for Y:", "RMSE for D:", "RMSE for Z:")
colnames(comp_tab) <- c("RF", "LASSO")
print(comp_tab, digits = 3)
```

# Weak Instruments?

```{r}
# using lfe package
summary(felm(dml2_lasso$dtil ~ dml2_lasso$ztil), robust = TRUE)
summary(felm(dml2_rf$dtil ~ dml2_rf$ztil), robust = TRUE)
```

## Anderson-Rubin Idea for Inference with Weak Instruments

As shown above, we may have weak instruments because the t-stat in the regression $\tilde D \sim \tilde Z$ is small relative to standard rules of thumb -- for example Stock and Yogo (2005) suggest accounting for weak instruments if the first stage F-statistic is less than 10 (and more recent work suggests even larger cutoffs).


 Here, we consider one specific approach (from Anderson and Rubin (1949)) to doing valid inference under weak identification based upon the statistic:
$$
C(\theta) = \frac{ |E_n [(\tilde Y -  \theta  \tilde D) \tilde Z]|^2}{ \mathbb{V}_n [(\tilde Y -  \theta \tilde  D) \tilde Z ]/n}.
$$
The empirical variance $\mathbb{V}_n$ is defined as:
\begin{align*}
\mathbb{V}_{n}[ g(W_i)] &:=   \mathbb{E}_{n}g(W_i)g(W_i)' - \mathbb{E}_{n}[ g(W_i)]\mathbb{E}_{n}[ g(W_i)]'.
\end{align*}

If $\theta_0 = \theta$, then $C(\theta) \overset{a}{\sim} N(0,1)^2 = \chi^2(1)$. Therefore,  we can reject the hypothesis $\theta_0 = \theta$ at level $a$ (for example $a = .05$ for a 5\% level test)  if $C(\theta)> c(1-a)$ where $c(1-a)$ is the $(1-a)$- quantile of a $\chi^2(1)$ variable.  The probability of falsely rejecting the true hypothesis is approximately $a \times 100\%$.  
To construct a $(1-a) \times 100$\% confidence region for $\theta$, we can then simply invert the test by collecting all parameter values that are not rejected at the $a$ level:
$$
CR(\theta) = \{ \theta \in \Theta: C(\theta)  \leq c(1-a)\}.
$$


In more complex settings with many controls or controls that enter with unknown functional form, we can simply replace the residuals $\tilde Y$, $\tilde D$, and $\tilde Z$ by machine
learned cross-fitted residuals $\check Y$, $\check D$, and $ \check Z$.  Thanks to the orthogonality of the IV moment condition underlying the formulation outlined above, we can formally assert that the properties of $C(\theta)$ and the subsequent testing procedure and confidence region for $\theta$ continue to hold when using cross-fitted residuals. We will further be able to apply the general procedure to cases where $D$
is a vector, with a suitable adjustment of the statistic $C(\theta)$.

```{r}
# DML-AR (DML with Anderson-Rubin)

dml_ar_pliv <- function(rY, rD, rZ, grid, alpha = .05) {
  n <- length(rY)
  Cstat <- rep(0, length(grid))
  for (i in seq_along(grid)) {
    Cstat[i] <- n * (mean((rY - grid[i] * rD) * rZ))^2 / var((rY - grid[i] * rD) * rZ)
  }
  LB <- min(grid[Cstat < qchisq(1 - alpha, 1)])
  UB <- max(grid[Cstat < qchisq(1 - alpha, 1)])
  plot(range(grid), range(c(Cstat)), type = "n", xlab = "Effect of institutions", ylab = "Statistic", main = " ")
  lines(grid, Cstat, lty = 1, col = 1)
  abline(h = qchisq(1 - alpha, 1), lty = 3, col = 4)
  abline(v = LB, lty = 3, col = 2)
  abline(v = UB, lty = 3, col = 2)
  return(list(UB = UB, LB = LB))
}
```

```{r}
dml_ar_pliv(
  rY = dml2_lasso$ytil, rD = dml2_lasso$dtil, rZ = dml2_lasso$ztil,
  grid = seq(-2, 2, by = .01)
)

dml_ar_pliv(
  rY = dml2_rf$ytil, rD = dml2_rf$dtil, rZ = dml2_rf$ztil,
  grid = seq(-2, 2, by = .01)
)
```

