---
title: An R Markdown document converted from "AC2/r-weak-iv-experiments.irnb"
output: html_document
---

# A Simple Example of Properties of IV estimator when Instruments are Weak

```{r}
install.packages("hdm")
```

```{r}
library(hdm)
```

Simulation Design

```{r}
# Simulation Design
set.seed(1)

n <- 100
beta <- .1 # .1 weak IV (change to 1.0 for strong IV)

# One realization
U <- rnorm(n)
Z <- rnorm(n) # generate instrument
D <- beta * Z + U # generate endogenougs variable
Y <- D + U # the true causal effect is 1

summary(lm(D ~ Z)) # first stage is very weak here when we set beta = .1
```

```{r}
summary(tsls(x = NULL, d = D, y = Y, z = Z))
```

Note that the instrument is weak here (strength of the instrument is controlled by setting $\beta$) -- the t-stat is smaller than any rule-of-thumb suggested in the literature (e.g. $\sqrt{10}$) .

# Run 10000 trials to evaluate distribution of the IV estimator

```{r}
# Simulation Design

set.seed(1)
B <- 10000 # trials
IVEst <- rep(0, B)

for (i in 1:B) {
  U <- rnorm(n)
  Z <- rnorm(n) # generate instrument
  D <- beta * Z + U # generate endogenougs variable
  Y <- D + U # the true causal effect is 1
  IVEst[i] <- coef(tsls(x = NULL, d = D, y = Y, z = Z))[1, 1]
}
```

# Plot the Actual Distribution against the Normal Approximation (based on Strong Instrument Assumption)

```{r}
plot(density(IVEst - 1, n = 1000, from = -5, to = 5),
  col = 4, xlim = c(-5, 5),
  xlab = "IV Estimator -True Effect", main = "Actual Distribution vs Gaussian"
)

val <- seq(-5, 5, by = .05)
var <- (1 / beta^2) * (1 / 100) # theoretical variance of IV
sd <- sqrt(var)
lines(val, dnorm(val, sd = sd), col = 2, lty = 2)

rejection_frequency <- sum((abs(IVEst - 1) / sd > 1.96)) / B

cat(c("Rejection Frequency is ", rejection_frequency, " while we expect it to be .05"))
```

