---
title: An R Markdown document converted from "CM3/r-dosearch.irnb"
output: html_document
---

# Dosearch for Causal Identification in DAGs.


This a simple notebook for teaching that illustrates capabilites of the "dosearch" package, which is a great tool.

NB. In my experience, the commands are sensitive to syntax ( e.g. spacing when -> are used), so be careful when changing to other examples.

```{r}
install.packages("dosearch")
```

```{r}
library("dosearch")
```

We start with the simplest graph, with the simplest example
where $D$ is policy, $Y$ is outcomes, $X$ is a confounder:
$$
D\to Y, \quad X \to (D,Y)
$$

Now suppose we want conditional average policy effect.

```{r}
data <- "p(y, d, x)" # data structure

query <- "p(y | do(d), x)" # query -- target parameter

graph <- "
  x -> y
  x -> d
  d -> y
"

dosearch(data, query, graph)
```

This recovers the correct identification formula for law of the counterfactual $Y(d)$ induced by $do(D=d)$:
$$
p_{Y(d)|X}(y|x) := p(y|do(d),x) = p(y|d,x).
$$

```{r}
data <- "p(y, d, x)"

query <- "p(y | do(d))"

graph <- "
  x -> y
  x -> d
  d -> y
"

dosearch(data, query, graph)
```

This recovers the correct identification formula:
$$
p_{Y(d)}(y) := p(y: do(d)) = \sum_{x}\left(p(x)p(y|d,x)\right)
$$
We integrate out $x$ in the previous formula.

Suppose we don't observe the confounder. The effect is generally not identified.

```{r}
data <- "p(y, d)"

query <- "p(y | do(d))"

graph <- "
  x -> y
  x -> d
  d -> y
"

dosearch(data, query, graph)
```

The next graph is an example of J. Pearl (different notation), where the graph is considerably more complicated. See "Pearl's Example" in the book - e.g. Figure 7.14. We are interested in $D \to Y$.

Here we try conditioning on $X_2$. This would block one backdoor path from $D$ to $Y$, but would open another path on which $X_2$ is a collider, so this shouldn't work. The application below gave a correct answer (after I put the spacings carefully).

```{r}
data <- "p(y, d, x2)" # observed only (Y, D, X_2)

query <- "p(y | do(d))" # target parameter

graph <- "
  z1 -> x1
  z1 -> x2
  z2 -> x2
  z2 -> x3
  x2 -> d
  x2 -> y
  x3 -> y
  x1 -> d
  d -> m
  m -> y
"

dosearch(data, query, graph)
```

Intuitively, we should add more common causes. For example, adding $X_3$ and using $S = (X_2, X_3)$ should work.

```{r}
data <- "p(y, d, x2, x3)"

conditional_query <- "p(y | do(d), x2, x3)" # can ID conditional average effect?
query <- "p(y | do(d))" # can ID unconditional effect?

graph <- "
  z1 -> x1
  z1 -> x2
  z2 -> x2
  z2 -> x3
  x2 -> d
  x2 -> y
  x3 -> y
  x1 -> d
  d -> m
  m -> y
"

print(dosearch(data, conditional_query, graph))
print(dosearch(data, query, graph))
```

This retrieves the correct formulas for counterfactual distributions of $Y(d)$ induced by $Do(D=d)$:

The conditional distribution is identified by
$$
p_{Y(d) \mid X_2, X_3}(y) := p(y |x_2, x_3: do(d)) = p(y|x_2,x_3,d).
$$

The unconditional distribution is obtained by integration out over $x_2$ and $x_3$:

$$
p_{Y(d) }(y) :=  p(y do(d)) = \sum_{x2,x3}\left(p(x_2,x_3)p(y|x_2,x_3,d)\right).
$$



Next we suppose that we observe only $(Y,D, M)$. Can we identify the effect $D \to Y$?  Can we use back-door-criterion twice to get $D \to M$ and $M \to Y$ affect? Yes, that's called front-door criterion -- so we just need to remember only the back-door and the fact that we can use it iteratively.

```{r}
data <- "p(y, d, m)"

query_dm <- "p(m | do(d))"
query_md <- "p(y | do(m))"
query <- "p(y | do(d))"

graph <- "
  z1 -> x1
  z1 -> x2
  z2 -> x2
  z2 -> x3
  x2 -> d
  x2 -> y
  x3 -> y
  x1 -> d
  d -> m
  m -> y
"
print(dosearch(data, query_dm, graph))
print(dosearch(data, query_md, graph))
print(dosearch(data, query, graph))
```

So we get identification results:
First,
$$
p_{M(d)}(m)  := p(m|do(d)) = p(m|d).
$$
Second,
$$
p_{Y(m)}(y) := p(y|do(m)) = \sum_{d}\left(p(d)p(y|d,m)\right),
$$
and the last by integrating the product of these two formulas:
$$
p_{Y(d)}(y) := p(y|do(d)) = \sum_{m}\left(p(m|d)\sum_{d}\left(p(d)p(y|d,m)\right)\right)
$$

The package is very rich and allows identification analysis, when the data comes from multiple sources. Suppose we observe marginal distributions $(Y,D)$  and $(D,M)$ only. Can we identify the effect of $D \to Y$. The answer is (guess) and the package correctly recovers it.

```{r}
data <- "p(y, m)
         p(m, d)"

query_dm <- "p(m | do(d))"
query_md <- "p(y | do(m))"
query <- "p(y | do(d))"

graph <- "
  z1 -> x1
  z1 -> x2
  z2 -> x2
  z2 -> x3
  x2 -> d
  x2 -> y
  x3 -> y
  x1 -> d
  d -> m
  m -> y
"
print(dosearch(data, query_dm, graph))
print(dosearch(data, query_md, graph))
print(dosearch(data, query, graph))
```

