---
title: An R Markdown document converted from "AC1/r-sensitivity-analysis-with-sensmakr-and-debiased-ml.irnb"
output: html_document
---

# Sensitivity Analysis for Unobserved Confounder with DML and Sensmakr

Here we experiment with using package "sensemakr" in conjunction with debiased ML.

## Partially Linear SEM

Consider the SEM
\begin{align}
Y :=~& \alpha D + \delta A + f_Y(X) + \epsilon_Y,  \\
D :=~& \gamma A + f_D(X) + \epsilon_D, \\
A :=~& f_A(X) + \epsilon_A, \\
X :=~&  \epsilon_X,
\end{align}
where, conditional on $X$, $\epsilon_Y, \epsilon_D, \epsilon_A$ are mean zero
and mutually uncorrelated. We further normalize
$$
E[\epsilon_A^2] =1.
$$
The key structural
parameter is $\alpha$: $$\alpha = \partial_d Y(d)$$
where $$Y(d) := (Y: do (D=d)).$$

To give context to our example, we can interpret $Y$ as earnings,
$D$ as education, $A$ as ability, and $X$ as a set of observed background variables. In this example, we can interpret $\alpha$ as the returns to schooling.

We start by applying the partialling out operator to get rid of the $X$'s in all of the equations. Define the partialling out operation of any random vector $V$ with respect to another random vector $X$ as the residual that is left after subtracting the best predictor of $V$ given $X$:
$$\tilde V = V - E [V \mid X].$$  
If $f$'s are linear, we can replace $E [V \mid X]$
by linear projection.  After partialling out, we have a simplified system:
\begin{align}
\tilde Y :=~& \alpha \tilde D + \delta \tilde A + \epsilon_Y,  \\
\tilde D :=~& \gamma \tilde A + \epsilon_D, \\
\tilde A :=~& \epsilon_A,
\end{align}
where $\epsilon_Y$, $\epsilon_D$, and $\epsilon_A$ are uncorrelated.

Then the projection of $\tilde Y$ on $\tilde D$ recovers
$$
\beta = E [\tilde Y \tilde D]/ E [\tilde D^2] = \alpha +  \phi,
$$
where
$$
\phi =  \delta \gamma/ E \left[(\gamma^2 + \epsilon^2_D)\right],
$$
is the omitted confounder bias or omitted variable bias.

The formula follows from inserting the expression for $\tilde D$ into the definition of $\beta$ and then simplifying the resulting expression using the assumptions on the $\epsilon$'s.

We can use this formula to bound $\phi$ directly by making assumptions on the size of $\delta$
and $\gamma$.  An alternative approach can be based on the following characterization,
based on partial $R^2$'s.  This characterization essentially follows
from Cinelli and Hazlett, with the slight difference that we have adapted
the result to the partially linear model.

*Theorem* [Omitted Confounder Bias in Terms of Partial $R^2$'s]

In the partially linear SEM setting above,
$$
\phi^2 = \frac{R^2_{\tilde Y \sim \tilde A \mid \tilde D} R^2_{\tilde D \sim \tilde A} }{ (1 - R^2_{\tilde D \sim \tilde A}) } \
\frac{E \left[ (\tilde Y - \beta \tilde D)^2 \right] }{E \left[ ( \tilde D )^2 \right]},
$$
where $R^2_{V \sim W \mid X}$ denotes the population $R^2$ in the linear regression of $V$ on $W$, after partialling out $X$ from $V$ and $W$ linearly.


Therefore, if we place bounds on how much of the variation in $\tilde Y$ and in $\tilde D$
the unobserved confounder $\tilde A$ is able to explain, we can bound the omitted confounder bias by $$\sqrt{\phi^2}.$$

# Empirical Example

We consider an empirical example based on data surrounding the Darfur war. Specifically, we are interested in the effect of having experienced direct war violence on attitudes towards peace. Data is described here
https://cran.r-project.org/web/packages/sensemakr/vignettes/sensemakr.html

The main outcome is attitude towards peace -- ``peacefactor``.
The key variable of interest is whether the responders were directly harmed (``directlyharmed``).
We want to know if being directly harmed in the conflict causes people to support peace-enforcing measures.
The measured confounders include dummy variables ``female``, ``farmer``, ``herder``, ``pastvoted``, along with ``age`` and household size (``hhsize``).
There is also a village indicator. We deal with village by and partialling out a full set of village dummy variables before conducting
the analysis. The standard errors will be clustered at the village level.


## Outline

We will:
- mimic the partialling out procedure with machine learning tools;
- invoke Sensmakr to compute $\phi^2$ and plot sensitivity results.

```{r}
# load package
install.packages("sensemakr")
install.packages("lfe")
install.packages("hdm")
install.packages("randomForest")
```

```{r}
library(sensemakr)
library(lfe)
library(hdm)
library(randomForest)

set.seed(1)
```

```{r}
file <- "https://raw.githubusercontent.com/CausalAIBook/MetricsMLNotebooks/main/data/darfur.csv"
data <- read.csv(file)
dim(data)
attach(darfur)
```

## Preprocessing
Take out village fixed effects and run basic linear analysis

```{r}
# get rid of village fixed effects
peacefactorR <- lm(peacefactor ~ village)$res
directlyharmedR <- lm(directlyharmed ~ village)$res
femaleR <- lm(female ~ village)$res
ageR <- lm(age ~ village)$res
farmerR <- lm(farmer_dar ~ village)$res
herderR <- lm(herder_dar ~ village)$res
pastvotedR <- lm(pastvoted ~ village)$res
hhsizeR <- lm(hhsize_darfur ~ village)$res


# Preliminary linear model analysis
summary(felm(peacefactorR ~ directlyharmedR + femaleR +
               ageR + farmerR + herderR + pastvotedR + hhsizeR | 0 | 0 | village))

# here we are clustering standard errors at the village level
summary(felm(peacefactorR ~ femaleR +
               ageR + farmerR + herderR + pastvotedR + hhsizeR | 0 | 0 | village))

summary(felm(directlyharmedR ~ femaleR +
               ageR + farmerR + herderR + pastvotedR + hhsizeR | 0 | 0 | village))
```

For benchmarking, we want to understand the "partial $R^2$" of the controls on the outcome after partialling out the variable of interest.

```{r}
# partial out variable of interest
peacefactorR2 <- lm(peacefactorR ~ directlyharmedR)$res
femaleR2 <- lm(femaleR ~ directlyharmedR)$res
ageR2 <- lm(ageR ~ directlyharmedR)$res
farmerR2 <- lm(farmerR ~ directlyharmedR)$res
herderR2 <- lm(herderR ~ directlyharmedR)$res
pastvotedR2 <- lm(pastvotedR ~ directlyharmedR)$res
hhsizeR2 <- lm(hhsizeR ~ directlyharmedR)$res


# R^2 of controls after partialling out variable of interest
summary(lm(peacefactorR2 ~ femaleR2 +
             ageR2 + farmerR2 + herderR2 + pastvotedR2 + hhsizeR2))
```

## Lasso for partialling out controls

```{r}
resY <- rlasso(peacefactorR ~ (femaleR + ageR + farmerR + herderR + pastvotedR + hhsizeR)^3,
               post = FALSE)$res

resD <- rlasso(directlyharmedR ~ (femaleR + ageR + farmerR + herderR + pastvotedR + hhsizeR)^3,
               post = FALSE)$res

print(c("Controls explain the following fraction of variance of Outcome", 1 - var(resY) / var(peacefactorR)))
print(c("Controls explain the following fraction of variance of Treatment", 1 - var(resD) / var(directlyharmedR)))

dml_darfur_model <- felm(resY ~ resD | 0 | 0 | village) # cluster SEs by village

summary(dml_darfur_model, robust = TRUE) # cluster SE by village

dml_darfur_model <- lm(resY ~ resD) # linear model to use as input in sensemakr
```

## Manual Bias Analysis

```{r}
# Main estimate
beta <- dml_darfur_model$coef[2]

# Hypothetical values of partial R2s
r2_yc <- .13
r2_dc <- .01

# Elements of the bias equation
kappa <- (r2_yc * r2_dc) / (1 - r2_dc)
varianceRatio <- mean(dml_darfur_model$res^2) / mean(resD^2)

# Compute square bias
BiasSq <- kappa * varianceRatio

# Compute absolute value of the bias
print(cat("absolute value of the bias:", sqrt(BiasSq)[1]))

# plotting
gridr2_dc <- seq(0, .3, by = .001)
gridr2_yc <- kappa * (1 - gridr2_dc) / gridr2_dc
gridr2_yc <- ifelse(gridr2_yc > 1, 1, gridr2_yc)
plot(gridr2_dc, gridr2_yc,
  type = "l", col = 4, xlab = "Partial R2 of Treatment with Confounder",
  ylab = "Partial R2 of Outcome with Confounder",
  main = paste("Combination of R2 such that |Bias| < ", round(sqrt(BiasSq), digits = 4))
)
```

## Bias Analysis with Sensemakr

```{r}
dml_darfur_sensitivity <- sensemakr(
  model = dml_darfur_model,
  treatment = "resD"
)
summary(dml_darfur_sensitivity)
plot(dml_darfur_sensitivity, nlevels = 15)
```

## Random Forest for partialling out

The following code does DML with clustered standard errors by village

```{r}
dml2_for_plm <- function(x, d, y, dreg, yreg, nfold = 2, clusterID) {
  nobs <- nrow(x) # number of observations
  foldid <- rep.int(1:nfold, times = ceiling(nobs / nfold))[sample.int(nobs)] # define folds indices
  I <- split(1:nobs, foldid) # split observation indices into folds
  ytil <- dtil <- rep(NA, nobs)
  cat("fold: ")
  for (b in seq_along(I)) {
    dfit <- dreg(x[-I[[b]], ], d[-I[[b]]]) # take a fold out
    yfit <- yreg(x[-I[[b]], ], y[-I[[b]]]) # take a foldt out
    dhat <- predict(dfit, x[I[[b]], ], type = "response") # predict the left-out fold
    yhat <- predict(yfit, x[I[[b]], ], type = "response") # predict the left-out fold
    dtil[I[[b]]] <- (d[I[[b]]] - dhat) # record residual for the left-out fold
    ytil[I[[b]]] <- (y[I[[b]]] - yhat) # record residial for the left-out fold
    cat(b, " ")
  }
  rfit <- felm(ytil ~ dtil | 0 | 0 | clusterID) # get clustered standard errors using felm
  rfitSummary <- summary(rfit)
  coef_est <- rfitSummary$coef[2] # extract coefficient
  se <- rfitSummary$coef[2, 2] # record robust standard error
  cat(sprintf("\ncoef (se) = %g (%g)\n", coef_est, se)) # printing output
  return(list(coef_est = coef_est, se = se, dtil = dtil, ytil = ytil)) # save output and residuals
}
```

```{r}
x <- model.matrix(~ femaleR + ageR + farmerR + herderR + pastvotedR + hhsizeR)
dim(x)

d <- directlyharmedR
y <- peacefactorR
# DML with Random Forest:
dreg <- function(x, d) {
  randomForest(x, d)
} # ML method=Forest
yreg <- function(x, y) {
  randomForest(x, y)
} # ML method=Forest

dml2_rf <- dml2_for_plm(x, d, y, dreg, yreg, nfold = 10, clusterID = village)

resY <- dml2_rf$ytil
resD <- dml2_rf$dtil

print(c("Controls explain the following fraction of variance of Outcome",
        max(1 - var(resY) / var(peacefactorR), 0)))
print(c("Controls explain the following fraction of variance of Treatment",
        max(1 - var(resD) / var(directlyharmedR), 0)))

dml_darfur_model <- lm(resY ~ resD)
```

## Bias Analysis with Sensemakr


```{r}
dml_darfur_sensitivity <- sensemakr(
  model = dml_darfur_model,
  treatment = "resD"
)
summary(dml_darfur_sensitivity)

plot(dml_darfur_sensitivity, nlevels = 15)
```

