---
title: An R Markdown document converted from "PM3/r_functional_approximation_by_nn_and_rf.irnb"
output: html_document
---

# Functional Approximations by Trees and Neural Networks

Here we show how the function
$$
x \mapsto exp(4 x)
$$
can be easily approximated by a tree-based methods (Trees, Random Forest) and a neural network (2 Layered Neural Network)

```{r}
install.packages("randomForest")
install.packages("rpart")
install.packages("gbm")
install.packages("keras")


library(randomForest)
library(rpart)
library(gbm)
library(keras)
```

# Functional Approximation by a Tree

We play around with the penalty level $cp$ below to illustrate how it affects the complexity of tree. Recall we may use this to prune the tree to improve predictive performance and lessen the noise in our final estimate. A simple penalty would be the number of leaves times a penalty level $\alpha$.

Specifics on the penalty can be found [here](https://cran.r-project.org/web/packages/rpart/rpart.pdf).

```{r}
set.seed(1)
X_train <- matrix(runif(1000),1000,1)
Y_train <- exp(4*X_train)  #Noiseless case  Y=g(X)
dim(X_train)


# shallow tree
TreeModel<- rpart(Y_train~X_train, cp=.01)  #cp is penalty level
pred.TM<- predict(TreeModel, newx=X_train)
plot(X_train, Y_train, type="p", pch=19, xlab="z", ylab="g(z)")
points(X_train, pred.TM, col=3, pch=19)
```

```{r}
set.seed(1)
X_train <- matrix(runif(1000),1000,1)
Y_train <- exp(4*X_train)  #Noiseless case  Y=g(X)
dim(X_train)


TreeModel<- rpart(Y_train~X_train, cp=.0005)  #cp is penalty level
pred.TM<- predict(TreeModel, newx=X_train)
plot(X_train, Y_train, type="p", pch=19, xlab="z", ylab="g(z)")
points(X_train, pred.TM, col=3, pch=19)
```

# Functional Approximation by RF

Here we show how the function
$$
x \mapsto exp(4 x)
$$
can be easily approximated by a tree-based method (Random Forest) and a neural network (2 Layered Neural Network)

```{r}
RFmodel<- randomForest(Y_train~X_train)
pred.RF<- predict(RFmodel, newdata=X_train)
plot(X_train, Y_train, type="p", pch=19, xlab="z", ylab="g(z)")
points(X_train, pred.RF, col=4,  pch=19,)
```

# Boosted Trees

```{r}
data_train = as.data.frame(cbind(X_train, Y_train))
BoostTreemodel<- gbm(Y_train~X_train, distribution= "gaussian",  n.trees=100, shrinkage=.01, interaction.depth
=3)
# shrinkage is "learning rate"
# n.trees is the number of boosting steps
# interaction.depth is the max depth of each tree
pred.BT<- predict(BoostTreemodel, newdata=data_train, n.trees=100)
plot(X_train, Y_train, type="p", pch=19, xlab="z", ylab="g(z)")
points(X_train, pred.BT, col=4,  pch=19,)
```

```{r}
data_train = as.data.frame(cbind(X_train, Y_train))
BoostTreemodel<- gbm(Y_train~X_train, distribution= "gaussian",  n.trees=1000, shrinkage=.01, interaction.depth
=3)
# shrinkage is "learning rate"
# n.trees is the number of boosting steps
# interaction.depth is the max depth of each tree
pred.BT<- predict(BoostTreemodel, newdata=data_train, n.trees=1000)
plot(X_train, Y_train, type="p", pch=19, xlab="z", ylab="g(z)")
points(X_train, pred.BT, col=4,  pch=19,)
```

# Same Example with a Neural Network

```{r}
build_model <- function() {
  model <- keras_model_sequential() %>%
    layer_dense(units = 200, activation = "relu",
                input_shape = 1)%>%
   layer_dense(units = 20, activation = "relu") %>%
    layer_dense(units = 1)

  model %>% compile(
    optimizer = optimizer_adam(lr = 0.01),
    loss = "mse",
    metrics = c("mae"),
  )
}
```

```{r}
model <- build_model()
summary(model)
```

```{r}
num_epochs <- 1
model %>% fit(X_train, Y_train,
                    epochs = num_epochs, batch_size = 10, verbose = 0)
pred.NN <- model %>% predict(X_train)
plot(X_train, Y_train, type="p", pch=19, xlab="z", ylab="g(z)")
points(X_train, pred.NN, col=4,  pch=19,)

```

```{r}
num_epochs <- 100
model %>% fit(X_train, Y_train,
                    epochs = num_epochs, batch_size = 10, verbose = 0)
pred.NN <- model %>% predict(X_train)
plot(X_train, Y_train, type="p", pch=19, xlab="z", ylab="g(z)")
points(X_train, pred.NN, col=4,  pch=19,)


```

### Using Early Stopping

```{r}
# Define the neural network architecture
model <- keras_model_sequential() %>%
  layer_dense(units = 200, activation = 'relu', input_shape = 1) %>%
  layer_dense(units = 20, activation = 'relu') %>%
  layer_dense(units = 1)  # Output layer with 1 unit for regression task

# Compile the model
model %>% compile(
  optimizer = optimizer_adam(lr = 0.01),
  loss = "mse",
  metrics = c("mae"),
)

summary(model)
```

```{r}
num_epochs <- 100

# Define early stopping based on validation set (20%) performance
early_stopping <- callback_early_stopping(monitor = 'val_loss', patience = 5)  # Patience set to 5 epochs (default in skorch is 5)

# Train the model
model %>% fit(
  X_train, Y_train,
  epochs = num_epochs,
  batch_size = 10,
  validation_split = 0.2,  # 20% validation set
  verbose = 0,
  callbacks = list(early_stopping)
)

pred.NN <- model %>% predict(X_train)
plot(X_train, Y_train, type="p", pch=19, xlab="z", ylab="g(z)")
points(X_train, pred.NN, col=4,  pch=19,)
```

