---
title: An R Markdown document converted from "PM3/r_ml_wage_prediction.irnb"
output: html_document
---

# Machine Learning Estimators for Wage Prediction

We illustrate how to predict an outcome variable Y in a high-dimensional setting, where the number of covariates $p$ is large in relation to the sample size $n$. So far we have used linear prediction rules, e.g. Lasso regression, for estimation.
Now, we also consider nonlinear prediction rules including tree-based methods.

```{r}
# Import relevant packages
install.packages("xtable")
install.packages("hdm")
install.packages("glmnet")
install.packages("randomForest")
install.packages("rpart")
install.packages("nnet")
install.packages("gbm")
install.packages("rpart.plot")
install.packages("keras")
```

```{r}
library(hdm)
library(xtable)
library(glmnet)
library(randomForest)
library(rpart)
library(nnet)
library(gbm)
library(rpart.plot)
library(keras)
```

## Data

Again, we consider data from the U.S. March Supplement of the Current Population Survey (CPS) in 2015.
The preproccessed sample consists of $5150$ never-married individuals.

```{r}
file <- "https://raw.githubusercontent.com/CausalAIBook/MetricsMLNotebooks/main/data/wage2015_subsample_inference.csv"
data <- read.csv(file)
dim(data)
```

The outcomes $Y_i$'s are hourly (log) wages of never-married workers living in the U.S. The raw regressors $Z_i$'s consist of a variety of characteristics, including experience, education and industry and occupation indicators.

```{r}
Z <- subset(data, select = -c(lwage, wage)) # regressors
colnames(Z)
```

The following figure shows the weekly wage distribution from the US survey data.

```{r}
hist(data$wage, xlab = "hourly wage", main = "Empirical wage distribution from the US survey data", breaks = 35)
```

Wages show a high degree of skewness. Hence, wages are transformed in almost all studies by
the logarithm.

## Analysis

Due to the skewness of the data, we are considering log wages which leads to the following regression model

$$log(wage) = g(Z) + \epsilon.$$

We will estimate the two sets of prediction rules: Linear and Nonlinear Models.
In linear models, we estimate the prediction rule of the form

$$\hat g(Z) = \hat \beta'X.$$
Again, we generate $X$ in two ways:

1. Basic Model:   $X$ consists of a set of raw regressors (e.g. gender, experience, education indicators, regional indicators).


2. Flexible Model:  $X$ consists of all raw regressors from the basic model plus occupation and industry indicators, transformations (e.g., ${exp}^2$ and ${exp}^3$) and additional two-way interactions.

To evaluate the out-of-sample performance, we split the data first.

```{r}
set.seed(1234)
training <- sample(nrow(data), nrow(data) * (3 / 4), replace = FALSE)

data_train <- data[training, ]
data_test <- data[-training, ]
```

We construct the two different model matrices $X_{basic}$ and $X_{flex}$ for both the training and the test sample:

```{r}
x_basic <- "sex + exp1 + shs + hsg+ scl + clg + mw + so + we + C(occ2)+ C(ind2)"
x_flex <- paste("sex + exp1 + shs + hsg + scl + clg + C(occ2) + C(ind2) + mw + so + we ",
                "+ (exp1 + exp2 + exp3 + exp4) * (shs + hsg + scl + clg + C(occ2) + C(ind2) + mw + so + we)")
formula_basic <- as.formula(paste("lwage", "~", x_basic))
formula_flex <- as.formula(paste("lwage", "~", x_flex))

model_x_basic_train <- model.matrix(formula_basic, data_train)
model_x_basic_test <- model.matrix(formula_basic, data_test)
p_basic <- dim(model_x_basic_train)[2]
model_x_flex_train <- model.matrix(formula_flex, data_train)
model_x_flex_test <- model.matrix(formula_flex, data_test)
p_flex <- dim(model_x_flex_train)[2]
```

```{r}
y_train <- data_train$lwage
y_test <- data_test$lwage
```

```{r}
p_basic
p_flex
```

As known from our first lab, the basic model consists of $51$ regressors and the flexible model of $246$ regressors. Let us fit our models to the training sample using the two different model specifications. We are starting by running a simple ols regression.

### OLS

We fit the basic model to our training data by running an ols regression and compute the mean squared error on the test sample.

```{r}
# ols (basic model)
fit_lm_basic <- lm(formula_basic, data_train)
```

```{r}
# Compute the Out-Of-Sample Performance
yhat_lm_basic <- predict(fit_lm_basic, newdata = data_test)
# MSE OLS (basic model)
cat("The mean squared error (MSE) using the basic model is equal to", mean((y_test - yhat_lm_basic)^2))
```

To determine the out-of-sample $MSE$ and the standard error in one step, we can use the function *lm*:

```{r}
mse_lm_basic <- summary(lm((y_test - yhat_lm_basic)^2 ~ 1))$coef[1:2]
mse_lm_basic
```

We also compute the out-of-sample $R^2$:

```{r}
r2_lm_basic <- 1 - mse_lm_basic[1] / var(y_test)
# MSE OLS (basic model)
cat("The R^2 using the basic model is equal to", r2_lm_basic)
```

We repeat the same procedure for the flexible model.

```{r}
# ols (flexible model)
fit_lm_flex <- lm(formula_flex, data_train)
# Compute the Out-Of-Sample Performance
options(warn = -1)
yhat_lm_flex <- predict(fit_lm_flex, newdata = data_test)
mse_lm_flex <- summary(lm((y_test - yhat_lm_flex)^2 ~ 1))$coef[1:2]
r2_lm_flex <- 1 - mse_lm_flex[1] / var(y_test)
cat("The R^2 using the flexible model is equal to", r2_lm_flex) # MSE OLS (flexible model)
```

We observe that ols regression works better for the basic model with smaller $p/n$ ratio. We now proceed by running lasso regressions and related penalized methods.

### Lasso, Ridge and Elastic Net

Considering the basic model, we run a lasso/post-lasso regression first and then we compute the measures for the out-of-sample performance. Note that applying the package *hdm* and the function *rlasso* we rely on a theoretical based choice of the penalty level $\lambda$ in the lasso regression.

```{r}
# lasso and variants
fit_rlasso <- hdm::rlasso(formula_basic, data_train, post = FALSE)
fit_rlasso_post <- hdm::rlasso(formula_basic, data_train, post = TRUE)
yhat_rlasso <- predict(fit_rlasso, newdata = data_test)
yhat_rlasso_post <- predict(fit_rlasso_post, newdata = data_test)

mse_lasso <- summary(lm((y_test - yhat_rlasso)^2 ~ 1))$coef[1:2]
mse_lasso_post <- summary(lm((y_test - yhat_rlasso_post)^2 ~ 1))$coef[1:2]

r2_lasso <- 1 - mse_lasso[1] / var(y_test)
r2_lasso_post <- 1 - mse_lasso_post[1] / var(y_test)
# R^2 lasso/post-lasso (basic model)
cat("The R^2 using the basic model is equal to", r2_lasso, "for lasso and", r2_lasso_post, "for post-lasso")
```

Now, we repeat the same procedure for the flexible model.

```{r}
fit_rlasso_flex <- hdm::rlasso(formula_flex, data_train, post = FALSE)
fit_rlasso_post_flex <- hdm::rlasso(formula_flex, data_train, post = TRUE)
yhat_rlasso_flex <- predict(fit_rlasso_flex, newdata = data_test)
yhat_rlasso_post_flex <- predict(fit_rlasso_post_flex, newdata = data_test)

mse_lasso_flex <- summary(lm((y_test - yhat_rlasso_flex)^2 ~ 1))$coef[1:2]
mse_lasso_post_flex <- summary(lm((y_test - yhat_rlasso_post_flex)^2 ~ 1))$coef[1:2]

# R^2 lasso/post-lasso (flexible model)
r2_lasso_flex <- 1 - mse_lasso_flex[1] / var(y_test)
r2_lasso_post_flex <- 1 - mse_lasso_post_flex[1] / var(y_test)
cat("The R^2 using the flexible model is equal to", r2_lasso_flex,
    "for lasso and", r2_lasso_post_flex, "for post-lasso")
```

In contrast to a theoretical based choice of the tuning parameter $\lambda$ in the lasso regression, we can also use cross-validation to determine the penalty level by applying the package *glmnet* and the function cv.glmnet. In this context, we also run a ridge and a elastic net regression by adjusting the parameter *alpha*.

```{r}
fit_lasso_cv <- cv.glmnet(model_x_basic_train, y_train, family = "gaussian", alpha = 1)
fit_ridge <- cv.glmnet(model_x_basic_train, y_train, family = "gaussian", alpha = 0)
fit_elnet <- cv.glmnet(model_x_basic_train, y_train, family = "gaussian", alpha = .5)

yhat_lasso_cv <- predict(fit_lasso_cv, newx = model_x_basic_test)
yhat_ridge <- predict(fit_ridge, newx = model_x_basic_test)
yhat_elnet <- predict(fit_elnet, newx = model_x_basic_test)

mse_lasso_cv <- summary(lm((y_test - yhat_lasso_cv)^2 ~ 1))$coef[1:2]
mse_ridge <- summary(lm((y_test - yhat_ridge)^2 ~ 1))$coef[1:2]
mse_elnet <- summary(lm((y_test - yhat_elnet)^2 ~ 1))$coef[1:2]

r2_lasso_cv <- 1 - mse_lasso_cv[1] / var(y_test)
r2_ridge <- 1 - mse_ridge[1] / var(y_test)
r2_elnet <- 1 - mse_elnet[1] / var(y_test)

# R^2 using cross-validation (basic model)
cat("R^2 using cross-validation for lasso, ridge and elastic net in the basic model:",
    r2_lasso_cv, r2_ridge, r2_elnet)
```

Note that the following calculations for the flexible model need some computation time.

```{r}
fit_lasso_cv_flex <- cv.glmnet(model_x_flex_train, y_train, family = "gaussian", alpha = 1)
fit_ridge_flex <- cv.glmnet(model_x_flex_train, y_train, family = "gaussian", alpha = 0)
fit_elnet_flex <- cv.glmnet(model_x_flex_train, y_train, family = "gaussian", alpha = .5)

yhat_lasso_cv_flex <- predict(fit_lasso_cv_flex, newx = model_x_flex_test)
yhat_ridge_flex <- predict(fit_ridge_flex, newx = model_x_flex_test)
yhat_elnet_flex <- predict(fit_elnet_flex, newx = model_x_flex_test)

mse_lasso_cv_flex <- summary(lm((y_test - yhat_lasso_cv_flex)^2 ~ 1))$coef[1:2]
mse_ridge_flex <- summary(lm((y_test - yhat_ridge_flex)^2 ~ 1))$coef[1:2]
mse_elnet_flex <- summary(lm((y_test - yhat_elnet_flex)^2 ~ 1))$coef[1:2]

r2_lasso_cv_flex <- 1 - mse_lasso_cv_flex[1] / var(y_test)
r2_ridge_flex <- 1 - mse_ridge_flex[1] / var(y_test)
r2_elnet_flex <- 1 - mse_elnet_flex[1] / var(y_test)

# R^2 using cross-validation (flexible model)
cat("R^2 using cross-validation for lasso, ridge and elastic net in the flexible model:",
    r2_lasso_cv_flex, r2_ridge_flex, r2_elnet_flex)
```

The performance of the lasso regression with cross-validated penalty is quite similar to the performance of lasso using a theoretical based choice of the tuning parameter.

#Non-linear models

Besides linear regression models, we consider nonlinear regression models to build a predictive model. We are applying regression trees, random forests, boosted trees and neural nets to estimate the regression function $g(X)$.

## Regression Trees

We fit a regression tree to the training data using the basic model. The variable *cp* controls the complexity of the regression tree, i.e. how deep we build the tree.

```{r}
# tree
fit_trees <- rpart(formula_basic, data_train, minbucket = 5, cp = 0.001)
# plotting the tree
prp(fit_trees, leaf.round = 1, space = 2, yspace = 2, split.space = 2, shadow.col = "gray", trace = 1)
```

An important method to improve predictive performance is called "Pruning the Tree". This
means the process of cutting down the branches of a tree. We apply pruning to the complex tree above to reduce the depth. Initially, we determine the optimal complexity of the regression tree.

```{r}
bestcp <- fit_trees$cptable[which.min(fit_trees$cptable[, "xerror"]), "CP"]
bestcp
```

Now, we can prune the tree and visualize the prediction rule.

```{r}
fit_prunedtree <- prune(fit_trees, cp = bestcp)
prp(fit_prunedtree, leaf.round = 1, space = 3, yspace = 3, split.space = 7,
    shadow.col = "gray", trace = 1, yesno = 1)
```

Finally, we calculate the mean-squared error and the $R^2$ on the test sample to evaluate the out-of-sample performance of the pruned tree.

```{r}
yhat_pt <- predict(fit_prunedtree, newdata = data_test)
mse_pt <- summary(lm((y_test - yhat_pt)^2 ~ 1))$coef[1:2]
r2_pt <- 1 - mse_pt[1] / var(y_test)

# R^2 of the pruned tree
cat("R^2 of the pruned tree:", r2_pt)
```

## Random Forest and Boosted Trees

In the next step, we apply the more advanced tree-based methods random forest and boosted trees.

```{r}
# random forest
fit_rf <- randomForest(model_x_basic_train, y_train, ntree = 2000, nodesize = 20, data = data_train)

## Evaluating the method
yhat_rf <- predict(fit_rf, newdata = model_x_basic_test) # prediction

mse_rf <- summary(lm((y_test - yhat_rf)^2 ~ 1))$coef[1:2]
r2_rf <- 1 - mse_rf[1] / var(y_test)
```

```{r}
# boosting
fit_boost <- gbm(formula_basic, data = data_train, distribution = "gaussian", bag.fraction = .5,
                 interaction.depth = 2, n.trees = 1000, shrinkage = .01)
best_boost <- gbm.perf(fit_boost, plot.it = FALSE) # cross-validation to determine when to stop

## Evaluating the method
yhat_boost <- predict(fit_boost, newdata = data_test, n.trees = best_boost)

mse_boost <- summary(lm((y_test - yhat_boost)^2 ~ 1))$coef[1:2]
r2_boost <- 1 - mse_boost[1] / var(y_test)
```

```{r}
# printing R^2
cat("R^2 of the random forest and boosted trees:", r2_rf, r2_boost)
```

## NNets

First, we need to determine the structure of our network. We are using the R package *keras* to build a simple sequential neural network with three dense layers -- 2 hidden and one output layer.

```{r}
# Define the neural network architecture
model <- keras_model_sequential() %>%
  layer_dense(units = 50, activation = "relu", input_shape = dim(model_x_basic_train)[2]) %>%
  layer_dense(units = 50, activation = "relu") %>%
  layer_dense(units = 1) # Output layer with 1 unit for regression task

# Compile the model
model %>% compile(
  optimizer = optimizer_adam(lr = 0.01),
  loss = "mse",
  metrics = c("mae"),
)

summary(model)
```

```{r}
num_epochs <- 100

# Define early stopping based on validation set (20%) performance
# Patience set to 5 epochs (default in skorch is 5)
early_stopping <- callback_early_stopping(monitor = "val_loss", patience = 5)

# Train the model
model %>% fit(
  model_x_basic_train, y_train,
  epochs = num_epochs,
  batch_size = 10,
  validation_split = 0.2, # 20% validation set
  verbose = 0,
  callbacks = list(early_stopping)
)
```

```{r}
# evaluating the performance
model %>% evaluate(model_x_basic_test, y_test, verbose = 0)
```

```{r}
# Calculating the performance measures
yhat_nn <- model %>% predict(model_x_basic_test)
mse_nn <- summary(lm((y_test - yhat_nn)^2 ~ 1))$coef[1:2]
r2_nn <- 1 - mse_nn[1] / var(y_test)
# printing R^2
cat("R^2 of the neural network:", r2_nn)
```

To conclude, let us have a look at our results.

## Results

```{r}
table <- matrix(0, 16, 3)
table[1, 1:2] <- mse_lm_basic
table[2, 1:2] <- mse_lm_flex
table[3, 1:2] <- mse_lasso
table[4, 1:2] <- mse_lasso_post
table[5, 1:2] <- mse_lasso_flex
table[6, 1:2] <- mse_lasso_post_flex
table[7, 1:2] <- mse_lasso_cv
table[8, 1:2] <- mse_ridge
table[9, 1:2] <- mse_elnet
table[10, 1:2] <- mse_lasso_cv_flex
table[11, 1:2] <- mse_ridge_flex
table[12, 1:2] <- mse_elnet_flex
table[13, 1:2] <- mse_rf
table[14, 1:2] <- mse_boost
table[15, 1:2] <- mse_pt
table[16, 1:2] <- mse_nn


table[1, 3] <- r2_lm_basic
table[2, 3] <- r2_lm_flex
table[3, 3] <- r2_lasso
table[4, 3] <- r2_lasso_post
table[5, 3] <- r2_lasso_flex
table[6, 3] <- r2_lasso_post_flex
table[7, 3] <- r2_lasso_cv
table[8, 3] <- r2_ridge
table[9, 3] <- r2_elnet
table[10, 3] <- r2_lasso_cv_flex
table[11, 3] <- r2_ridge_flex
table[12, 3] <- r2_elnet_flex
table[13, 3] <- r2_rf
table[14, 3] <- r2_boost
table[15, 3] <- r2_pt
table[16, 3] <- r2_nn


colnames(table) <- c("MSE", "S.E. for MSE", "R-squared")
rownames(table) <- c(
  "Least Squares (basic)", "Least Squares (flexible)", "Lasso", "Post-Lasso",
  "Lasso (flexible)", "Post-Lasso (flexible)",
  "Cross-Validated lasso", "Cross-Validated ridge", "Cross-Validated elnet",
  "Cross-Validated lasso (flexible)", "Cross-Validated ridge (flexible)", "Cross-Validated elnet (flexible)",
  "Random Forest", "Boosted Trees", "Pruned Tree", "Neural Net (Early)"
)
tab <- xtable(table, digits = 3)
print(tab, type = "latex") # set type="latex" for printing table in LaTeX
tab
```

Above, we displayed the results for a single split of data into the training and testing part. The table shows the test MSE in column 1 as well as the standard error in column 2 and the test $R^2$
in column 3. We see that most models perform similarly. For most of these methods, test MSEs are within one standard error of each other. Remarkably, OLS with just the basic variables performs extremely well. However, OLS on a flexible model with many regressors performs very poorly giving nearly the highest test MSE. It is worth noticing that, as this is just a simple illustration meant to be relatively quick, the nonlinear models are not tuned. Thus, there is potential to improve the performance of the nonlinear methods we used in the analysis.

### Ensemble learning

In the final step, we can build a prediction model by combing the strengths of the models we considered so far. This ensemble method is of the form
	$$ f(x) = \sum_{k=1}^K \alpha_k f_k(x) $$
where the $f_k$'s denote our prediction rules from the table above and the $\alpha_k$'s are the corresponding weights.

We first estimate the weights without penalization.

```{r}
ensemble_ols <- summary(lm(y_test ~ yhat_lm_basic + yhat_lm_flex + yhat_rlasso + yhat_rlasso_flex +
                             yhat_rlasso_post + yhat_rlasso_post_flex + yhat_lasso_cv + yhat_lasso_cv_flex +
                             yhat_ridge + yhat_ridge_flex + yhat_elnet + yhat_elnet_flex +
                             yhat_pt + yhat_rf + yhat_boost + yhat_nn))
ensemble_ols
```

Alternatively, we can determine the weights via lasso regression.

```{r}
ensemble_lasso <- summary(hdm::rlasso(y_test ~ yhat_lm_basic + yhat_lm_flex + yhat_rlasso + yhat_rlasso_flex +
                                        yhat_rlasso_post + yhat_rlasso_post_flex + yhat_lasso_cv + yhat_lasso_cv_flex +
                                        yhat_ridge + yhat_ridge_flex + yhat_elnet + yhat_elnet_flex +
                                        yhat_pt + yhat_rf + yhat_boost + yhat_nn))
ensemble_lasso
```

The estimated weights are shown in the following table.

```{r}
table <- matrix(0, 17, 2)
table[1:17, 1] <- ensemble_ols$coef[1:17]
table[1:17, 2] <- ensemble_lasso$coef[1:17]

colnames(table) <- c("Weight OLS", "Weight Lasso")

rownames(table) <- c(
  "Constant", "Least Squares (basic)", "Least Squares (flexible)", "Lasso (basic)",
  "Lasso (flexible)", "Post-Lasso (basic)", "Post-Lasso (flexible)", "LassoCV (basic)",
  "Lasso CV (flexible)", "Ridge CV (basic)", "Ridge CV (flexible)", "ElNet CV (basic)",
  "ElNet CV (flexible)", "Pruned Tree", "Random Forest", "Boosted Trees", "Neural Net"
)
tab <- xtable(table, digits = 3)
print(tab, type = "latex") # set type="latex" for printing table in LaTeX
tab
```

We note the superior $R^2$ performance of the ensembles. Though for more unbiased performance evaluation, we should have left out a third sample to validate the performance of the stacked model.

```{r}
# print ensemble R^2
cat("R^2 of stacking with LS weights:", ensemble_ols$adj.r.squared, "\n")
cat("R^2 of stacking with Lasso weights:", ensemble_lasso$adj.r.squared, "\n")
```

# Automatic Machine Learning with H20 AutoML

We illustrate how to predict an outcome variable Y in a high-dimensional setting, using the AutoML package *H2O* that covers the complete pipeline from the raw dataset to the deployable machine learning model. In last few years, AutoML or automated machine learning has become widely popular among data science community. Again, we reanalyze the wage prediction problem using data from the U.S. March Supplement of the Current Population Survey (CPS) in 2015.

We can use AutoML as a benchmark and compare it to the methods that we used previously where we applied one machine learning method after the other.

On a unix system, installation of the package h2o also requires the RCurl package, which requires the libcurl library to have been installed. If the installation of the h2o package fails, try installing the libcurl package first and repeating the cell. You can install the libcurl package for instance by running:
```bash
sudo apt-get install -y libcurl4-openssl-dev
```

```{r}
# load the H2O package
install.packages("h2o")
```

```{r}
library(h2o)
```

```{r}
# start h2o cluster
h2o.init()
```

```{r}
# convert data as h2o type
train_h <- as.h2o(data_train)
test_h <- as.h2o(data_test)

# have a look at the data
h2o.describe(train_h)
```

```{r}
y_name <- "lwage"
x_names <- setdiff(names(data), c("lwage", "wage", "occ", "ind"))

# run AutoML for 10 base models and a maximal runtime of 100 seconds
aml <- h2o.automl(
  x = x_names, y = y_name,
  training_frame = train_h,
  leaderboard_frame = test_h,
  max_models = 10,
  seed = 1,
  max_runtime_secs = 100
)
# AutoML Leaderboard
lb <- aml@leaderboard
print(lb, n = nrow(lb))
```

We see that two Stacked Ensembles are at the top of the leaderboard. Stacked Ensembles often outperform a single model. The out-of-sample (test) MSE of the leading model is given by

```{r}
aml@leaderboard$mse[1]
```

The in-sample performance can be evaluated by

```{r}
aml@leader
```

This is in line with our previous results. To understand how the ensemble works, let's take a peek inside the Stacked Ensemble "All Models" model.  The "All Models" ensemble is an ensemble of all of the individual models in the AutoML run.  This is often the top performing model on the leaderboard.

```{r}
model_ids <- as.data.frame(aml@leaderboard$model_id)[, 1]
# Get the "All Models" Stacked Ensemble model
se <- h2o.getModel(grep("StackedEnsemble_AllModels", model_ids, value = TRUE)[1])
# Get the Stacked Ensemble metalearner model
metalearner <- se@model$metalearner_model
h2o.varimp(metalearner)
```

The table above gives us the variable importance of the metalearner in the ensemble. The AutoML Stacked Ensembles use the default metalearner algorithm (GLM with non-negative weights), so the variable importance of the metalearner is actually the standardized coefficient magnitudes of the GLM.

```{r}
h2o.varimp_plot(metalearner)
```

## Generating Predictions Using Leader Model

We can also generate predictions on a test sample using the leader model object.

```{r}
pred <- as.matrix(h2o.predict(aml@leader, test_h)) # make prediction using x data from the test sample
head(pred)
```

```{r}
y_test <- as.matrix(test_h$lwage)
r2_test <- 1 - summary(lm((y_test - pred)^2 ~ 1))$coef[1] / var(y_test)
cat("MSE, SE, R^2:", summary(lm((y_test - pred)^2 ~ 1))$coef[1:2], r2_test)
```

We observe both a similar MSE and $R^2$ relative to the better performing models in our previous results.

```{r}
h2o.shutdown(prompt = FALSE) # shut down the h20 automatically without prompting user
```

