---
title: An R Markdown document converted from "PM4/r-dml-401k.irnb"
output: html_document
---

# Inference on Predictive and Causal Effects in High-Dimensional Nonlinear Models

## Impact of 401(k) on  Financial Wealth

As a practical illustration of the methods developed in this lecture, we consider estimation of the effect of 401(k) eligibility and participation
on accumulated assets. 401(k) plans are pension accounts sponsored by employers. The key problem in determining the effect of participation in 401(k) plans on accumulated assets is saver heterogeneity coupled with the fact that the decision to enroll in a 401(k) is non-random. It is generally recognized that some people have a higher preference for saving than others. It also seems likely that those individuals with high unobserved preference for saving would be most likely to choose to participate in tax-advantaged retirement savings plans and would tend to have otherwise high amounts of accumulated assets. The presence of unobserved savings preferences with these properties then implies that conventional estimates that do not account for saver heterogeneity and endogeneity of participation will be biased upward, tending to overstate the savings effects of 401(k) participation.

One can argue that eligibility for enrolling in a 401(k) plan in this data can be taken as exogenous after conditioning on a few observables of which the most important for their argument is income. The basic idea is that, at least around the time 401(k)’s initially became available, people were unlikely to be basing their employment decisions on whether an employer offered a 401(k) but would instead focus on income and other aspects of the job.

```{r}
install.packages("xtable")
install.packages("hdm")
install.packages("sandwich")
install.packages("ggplot2")
install.packages("randomForest")
install.packages("glmnet")
install.packages("rpart")
install.packages("gbm")
```

```{r}
library(xtable)
library(hdm)
library(sandwich)
library(ggplot2)
library(randomForest)
library(data.table)
library(glmnet)
library(rpart)
library(gbm)
```

### Data

The raw dataset can be found [here](https://raw.githubusercontent.com/CausalAIBook/MetricsMLNotebooks/main/data/401k.csv).
The data set can be loaded from the `hdm` package for R directly by typing:


```{r}
data(pension)
data <- pension
dim(data)
```

See the "Details" section on the description of the data set, which can be accessed by

```{r}
help(pension)
```

The data consist of 9,915 observations at the household level drawn from the 1991 Survey of Income and Program Participation (SIPP).  All the variables are referred to 1990. We use net financial assets (*net\_tfa*) as the outcome variable, $Y$,  in our analysis. The net financial assets are computed as the sum of IRA balances, 401(k) balances, checking accounts, saving bonds, other interest-earning accounts, other interest-earning assets, stocks, and mutual funds less non mortgage debts.

Among the $9915$ individuals, $3682$ are eligible to participate in the program. The variable *e401* indicates eligibility and *p401* indicates participation, respectively.

```{r}
hist_e401 <- ggplot(data, aes(x = e401, fill = factor(e401))) +
  geom_bar()
hist_e401
```

Eligibility is highly associated with financial wealth:

```{r}
dens_net_tfa <- ggplot(data, aes(x = net_tfa, color = factor(e401), fill = factor(e401))) +
  geom_density() +
  xlim(c(-20000, 150000)) +
  facet_wrap(. ~ e401)

dens_net_tfa
```

The unconditional APE of e401 is about $19559$:

```{r}
e1 <- data[data$e401 == 1, ]
e0 <- data[data$e401 == 0, ]
round(mean(e1$net_tfa) - mean(e0$net_tfa), 0)
```

Among the $3682$ individuals that  are eligible, $2594$ decided to participate in the program. The unconditional APE of p401 is about $27372$:

```{r}
p1 <- data[data$p401 == 1, ]
p0 <- data[data$p401 == 0, ]
round(mean(p1$net_tfa) - mean(p0$net_tfa), 0)
```

As discussed, these estimates are biased since they do not account for saver heterogeneity and endogeneity of participation.

```{r}
# outcome variable
y <- data[, "net_tfa"]
# treatment variable
D <- data[, "e401"]
D2 <- data[, "p401"]
D3 <- data[, "a401"]

columns_to_drop <- c(
  "e401", "p401", "a401", "tw", "tfa", "net_tfa", "tfa_he",
  "hval", "hmort", "hequity",
  "nifa", "net_nifa", "net_n401", "ira",
  "dum91", "icat", "ecat", "zhat",
  "i1", "i2", "i3", "i4", "i5", "i6", "i7",
  "a1", "a2", "a3", "a4", "a5"
)

# covariates
X <- data[, !(names(data) %in% columns_to_drop)]
```

```{r}
# Constructing the controls
x_formula <- paste("~ 0 + poly(age, 6, raw=TRUE) + poly(inc, 8, raw=TRUE) + poly(educ, 4, raw=TRUE) ",
                   "+ poly(fsize, 2, raw=TRUE) + male + marr + twoearn + db + pira + hown")
X <- as.data.table(model.frame(x_formula, X))
head(X)
```

## Estimating the ATE of 401(k) Eligibility on Net Financial Assets

We are interested in valid estimators of the average treatment effect of `e401` and `p401` on `net_tfa`. We start using ML approaches to estimate the function $g_0$ and $m_0$ in the following PLR model:

\begin{align}
 &  Y = D\theta_0 + g_0(X) + \zeta,  &  E[\zeta \mid D,X]= 0,\\
 & D = m_0(X) +  V,   &  E[V \mid X] = 0.
\end{align}

## Partially Linear Regression Models (PLR)

```{r}
dml2_for_plm <- function(x, d, y, dreg, yreg, nfold = 3, method = "regression") {
  nobs <- nrow(x) # number of observations
  foldid <- rep.int(1:nfold, times = ceiling(nobs / nfold))[sample.int(nobs)] # define folds indices
  I <- split(1:nobs, foldid) # split observation indices into folds
  ytil <- dtil <- rep(NA, nobs)
  cat("fold: ")
  for (b in seq_along(I)) {
    if (method == "regression") {
      dfit <- dreg(x[-I[[b]], ], d[-I[[b]]]) # take a fold out
      yfit <- yreg(x[-I[[b]], ], y[-I[[b]]]) # take a foldt out
      dhat <- predict(dfit, x[I[[b]], ], type = "response") # predict the left-out fold
      yhat <- predict(yfit, x[I[[b]], ], type = "response") # predict the left-out fold
      dtil[I[[b]]] <- (d[I[[b]]] - dhat) # record residual for the left-out fold
      ytil[I[[b]]] <- (y[I[[b]]] - yhat) # record residial for the left-out fold
    } else if (method == "randomforest") {
      dfit <- dreg(x[-I[[b]], ], as.factor(d)[-I[[b]]]) # take a fold out
      yfit <- yreg(x[-I[[b]], ], y[-I[[b]]]) # take a fold out
      dhat <- predict(dfit, x[I[[b]], ], type = "prob")[, 2] # predict the left-out fold
      yhat <- predict(yfit, x[I[[b]], ], type = "response") # predict the left-out fold
      dtil[I[[b]]] <- (d[I[[b]]] - dhat) # record residual for the left-out fold
      ytil[I[[b]]] <- (y[I[[b]]] - yhat) # record residial for the left-out fold
    } else if (method == "decisiontrees") {
      dfit <- dreg(x[-I[[b]], ], as.factor(d)[-I[[b]]]) # take a fold out
      yfit <- yreg(x[-I[[b]], ], y[-I[[b]]]) # take a fold out
      dhat <- predict(dfit, x[I[[b]], ])[, 2] # predict the left-out fold
      yhat <- predict(yfit, x[I[[b]], ]) # predict the left-out fold
      dtil[I[[b]]] <- (d[I[[b]]] - dhat) # record residual for the left-out fold
      ytil[I[[b]]] <- (y[I[[b]]] - yhat) # record residial for the left-out fold
    } else if (method == "boostedtrees") {
      dfit <- dreg(x[-I[[b]], ], d[-I[[b]]]) # take a fold out
      yfit <- yreg(x[-I[[b]], ], y[-I[[b]]]) # take a fold out
      dhat <- predict(dfit, x[I[[b]], ], type = "response") # predict the left-out fold
      yhat <- predict(yfit, x[I[[b]], ], type = "response") # predict the left-out fold
      dtil[I[[b]]] <- (d[I[[b]]] - dhat) # record residual for the left-out fold
      ytil[I[[b]]] <- (y[I[[b]]] - yhat) # record residial for the left-out fold
    }
    cat(b, " ")
  }
  rfit <- lm(ytil ~ dtil) # estimate the main parameter by regressing one residual on the other
  coef_est <- coef(rfit)[2] # extract coefficient
  se <- sqrt(vcovHC(rfit)[2, 2]) # record robust standard error
  cat(sprintf("\ncoef (se) = %g (%g)\n", coef_est, se)) # printing output
  return(list(coef_est = coef_est, se = se, dtil = dtil, ytil = ytil)) # save output and residuals
}
```

```{r}
summaryPLR <- function(point, stderr, resD, resy, name) {
  data <- data.frame(
    estimate = point, # point estimate
    stderr = stderr, # standard error
    lower = point - 1.96 * stderr, # lower end of 95% confidence interval
    upper = point + 1.96 * stderr, # upper end of 95% confidence interval
    `rmse y` = sqrt(mean(resy^2)), # RMSE of model that predicts outcome y
    `rmse D` = sqrt(mean(resD^2)), # RMSE of model that predicts treatment D
    `accuracy D` = mean(abs(resD) < 0.5) # binary classification accuracy of model for D
  )
  rownames(data) <- name
  return(data)
}
```

#### Double Lasso with Cross-Fitting

```{r}
# DML with LassoCV
set.seed(123)
cat(sprintf("\nDML with Lasso CV \n"))

dreg_lasso_cv <- function(x, d) {
  cv.glmnet(x, d, family = "gaussian", alpha = 1, nfolds = 5)
}
yreg_lasso_cv <- function(x, y) {
  cv.glmnet(x, y, family = "gaussian", alpha = 1, nfolds = 5)
}

dml2_results <- dml2_for_plm(as.matrix(X), D, y, dreg_lasso_cv, yreg_lasso_cv, nfold = 5)

sum_lasso_cv <- summaryPLR(dml2_results$coef_est, dml2_results$se, dml2_results$dtil,
                           dml2_results$ytil, name = "LassoCV")
tableplr <- data.frame()
tableplr <- rbind(sum_lasso_cv)
tableplr
```

```{r}
# Because residuals are output, reconstruct fitted values for use in ensemble
dhat_lasso <- D - dml2_results$dtil
yhat_lasso <- y - dml2_results$ytil
```

#### Using a $\ell_2$ Penalized Logistic Regression for D

Note we are using the $\ell_2$ penalty here. You can use the $\ell_1$ penalty as well, but computation will take longer.

```{r}
# DML with Lasso/Logistic
set.seed(123)
cat(sprintf("\nDML with Lasso/Logistic \n"))

dreg_logistic_cv <- function(x, d) {
  cv.glmnet(x, d, family = "binomial", alpha = 0, nfolds = 5)
}
yreg_lasso_cv <- function(x, y) {
  cv.glmnet(x, y, family = "gaussian", alpha = 1, nfolds = 5)
}

dml2_results <- dml2_for_plm(as.matrix(X), D, y, dreg_logistic_cv, yreg_lasso_cv, nfold = 5)
sum_lasso_logistic_cv <- summaryPLR(dml2_results$coef_est, dml2_results$se, dml2_results$dtil,
                                    dml2_results$ytil, name = "LassoCV/LogisticCV")
tableplr <- rbind(tableplr, sum_lasso_logistic_cv)
tableplr
```

```{r}
# Because residuals are output, reconstruct fitted values for use in ensemble
dhat_lasso_logistic <- D - dml2_results$dtil
yhat_lasso_logistic <- y - dml2_results$ytil
```

#### Random Forests

```{r}
# DML with Random Forest
set.seed(123)
cat(sprintf("\nDML with Random Forest \n"))

dreg_rf <- function(x, d) {
  randomForest(x, d, ntree = 1000, nodesize = 10)
} # ML method=Forest
yreg_rf <- function(x, y) {
  randomForest(x, y, ntree = 1000, nodesize = 10)
} # ML method=Forest

dml2_results <- dml2_for_plm(as.matrix(X), D, y, dreg_rf, yreg_rf, nfold = 5, method = "randomforest")
sum_rf <- summaryPLR(dml2_results$coef_est, dml2_results$se, dml2_results$dtil,
                     dml2_results$ytil, name = "Random Forest")
tableplr <- rbind(tableplr, sum_rf)
tableplr
```

```{r}
# Because residuals are output, reconstruct fitted values for use in ensemble
dhat_rf <- D - dml2_results$dtil
dhat_rf <- y - dml2_results$ytil
```

#### Decision Trees

```{r}
# DML with Decision Trees
set.seed(123)
cat(sprintf("\nDML with Decision Trees \n"))

dreg_tr <- function(x, d) {
  rpart(as.formula("D~."), cbind(data.frame(D = d), x), method = "class", minbucket = 10, cp = 0.001)
}
dreg_tr <- function(x, y) {
  rpart(as.formula("y~."), cbind(data.frame(y = y), x), minbucket = 10, cp = 0.001)
}

# decision tree takes in X as dataframe, not matrix/array
dml2_results <- dml2_for_plm(X, D, y, dreg_tr, dreg_tr, nfold = 5, method = "decisiontrees")
sum_tr <- summaryPLR(dml2_results$coef_est, dml2_results$se, dml2_results$dtil,
                     dml2_results$ytil, name = "Decision Trees")
tableplr <- rbind(tableplr, sum_tr)
tableplr
```

```{r}
# Because residuals are output, reconstruct fitted values for use in ensemble
dhat_tr <- D - dml2_results$dtil
yhat_tr <- y - dml2_results$ytil
```


Ideally, we would do (semi) cross-fitting with AutoML in order to find good first-stage models and re-run DML with these models. Unfortunately this is not easy to do in R. In the case of semi cross-fitting, we can use R's H20 AutoML trained on the entire training set $y\sim X$, $D \sim X$, $Z\sim X$ to determine the best model (eg ensemble), but H20 does not allow you to extract said model so we can re-use that in DML.

#### Boosted Trees

```{r}
# DML with Boosted Trees
set.seed(123)
cat(sprintf("\nDML with Boosted Trees \n"))

# NB: early stopping cannot easily be implemented with gbm
## set n.trees = best, where best <- gbm.perf(dreg_boost, plot.it = FALSE)
dreg_boost <- function(x, d) {
  gbm(as.formula("D~."), cbind(data.frame(D = d), x), distribution = "bernoulli",
      interaction.depth = 2, n.trees = 100, shrinkage = .1)
}
yreg_boost <- function(x, y) {
  gbm(as.formula("y~."), cbind(data.frame(y = y), x), distribution = "gaussian",
      interaction.depth = 2, n.trees = 100, shrinkage = .1)
}

# passing these through regression as type="response", and D should not be factor!
dml2_results <- dml2_for_plm(X, D, y, dreg_boost, yreg_boost, nfold = 5, method = "boostedtrees")
sum_boost <- summaryPLR(dml2_results$coef_est, dml2_results$se, dml2_results$dtil,
                        dml2_results$ytil, name = "Boosted Trees")
tableplr <- rbind(tableplr, sum_boost)
tableplr
```

```{r}
# Because residuals are output, reconstruct fitted values for use in ensemble
dhat_boost <- D - dml2_results$dtil
yhat_boost <- y - dml2_results$ytil
```

## Ensembles

Boosted trees give the best RMSE for both Y and D, so the ensemble based on choosing the best performing prediction rule is identical to boosting in this case.

```{r}
# Best fit is boosted trees for both D and Y

sum_best <- summaryPLR(dml2_results$coef_est, dml2_results$se, dml2_results$dtil,
                       dml2_results$ytil, name = "Best")
tableplr <- rbind(tableplr, sum_best)
tableplr
```

We'll form a model average with unconstrained least squares weights.

```{r}
# Least squares model average

ma_dtil <- lm(D ~ dhat_lasso + dhat_lasso_logistic + dhat_rf + dhat_tr + dhat_boost)$residuals
ma_ytil <- lm(y ~ yhat_lasso + yhat_lasso_logistic + dhat_rf + yhat_tr + yhat_boost)$residuals

rfit <- lm(ma_ytil ~ ma_dtil) # estimate the main parameter by regressing one residual on the other
coef_est <- coef(rfit)[2] # extract coefficient
se <- sqrt(vcovHC(rfit)[2, 2]) # record robust standard error

sum.ma <- summaryPLR(coef_est, se, ma_dtil, ma_ytil, name = "Model Average")
tableplr <- rbind(tableplr, sum.ma)
tableplr
```

## Interactive Regression Model (IRM)

Next, we consider estimation of average treatment effects when treatment effects are fully heterogeneous:

 \begin{align}
 & Y  = g_0(D, X) + U,  &  \quad E[U \mid X, D]= 0,\\
  & D  = m_0(X) + V,  & \quad  E[V\mid X] = 0.
\end{align}

To reduce the disproportionate impact of extreme propensity score weights in the interactive model
we trim the propensity scores which are close to the bounds.

```{r}
dml2_for_irm <- function(x, d, y, dreg, yreg0, yreg1, trimming = 0.01, nfold = 5, method = "regression") {
  yhat0 <- rep(0, length(y))
  yhat1 <- rep(0, length(y))
  Dhat <- rep(0, length(d))

  nobs <- nrow(x) # number of observations
  foldid <- rep.int(1:nfold, times = ceiling(nobs / nfold))[sample.int(nobs)] # define folds indices
  I <- split(1:nobs, foldid) # split observation indices into folds
  ytil <- dtil <- rep(NA, nobs)

  cat("fold: ")
  for (b in seq_along(I)) {
    # define helpful variables
    Dnotb <- d[-I[[b]]]
    Xb <- X[I[[b]], ]
    Xnotb <- X[-I[[b]], ]

    # training dfs subsetted on the -I[[b]] fold
    XD0 <- X[-I[[b]], ][d[-I[[b]]] == 0]
    yD0 <- y[-I[[b]]][d[-I[[b]]] == 0]
    XD1 <- X[-I[[b]], ][d[-I[[b]]] == 1]
    yD1 <- y[-I[[b]]][d[-I[[b]]] == 1]

    if (method == "regression") {
      yfit0 <- yreg0(as.matrix(XD0), yD0)
      yfit1 <- yreg1(as.matrix(XD1), yD1)
      yhat0[I[[b]]] <- predict(yfit0, as.matrix(Xb)) # default is type = "response" for glmnet family gaussian
      yhat1[I[[b]]] <- predict(yfit1, as.matrix(Xb))
    } else if (method == "randomforest") {
      yfit0 <- yreg0(XD0, yD0)
      yfit1 <- yreg1(XD1, yD1)
      yhat0[I[[b]]] <- predict(yfit0, Xb) # default is type = "response" for rf
      yhat1[I[[b]]] <- predict(yfit1, Xb)
    } else if (method == "decisiontrees") {
      yfit0 <- yreg0(XD0, yD0)
      yfit1 <- yreg1(XD1, yD1)
      yhat0[I[[b]]] <- predict(yfit0, Xb) # default is type = "vector" for decision
      yhat1[I[[b]]] <- predict(yfit1, Xb)
    } else if (method == "boostedtrees") {
      yfit0 <- yreg0(as.data.frame(XD0), yD0)
      yfit1 <- yreg1(as.data.frame(XD1), yD1)
      yhat0[I[[b]]] <- predict(yfit0, Xb) # default is type = "response" for boosted
      yhat1[I[[b]]] <- predict(yfit1, Xb)
    }

    # propensity scores:
    if (method == "regression") {
      dfit_b <- dreg(as.matrix(Xnotb), Dnotb)
      dhat_b <- predict(dfit_b, as.matrix(Xb), type = "response") # default is type="link" for family binomial!
    } else if (method == "randomforest") {
      dfit_b <- dreg(Xnotb, as.factor(Dnotb))
      dhat_b <- predict(dfit_b, Xb, type = "prob")[, 2]
    } else if (method == "decisiontrees") {
      dfit_b <- dreg(Xnotb, Dnotb)
      dhat_b <- predict(dfit_b, Xb)[, 2]
    } else if (method == "boostedtrees") {
      dfit_b <- dreg(as.data.frame(Xnotb), Dnotb)
      dhat_b <- predict(dfit_b, Xb, type = "response")
    }
    dhat_b <- pmax(pmin(dhat_b, 1 - trimming), trimming) # trimming so scores are between [trimming, (1-trimming)]
    Dhat[I[[b]]] <- dhat_b

    cat(b, " ")
  }

  # Prediction of treatment and outcome for observed instrument
  yhat <- yhat0 * (1 - D) + yhat1 * D
  # residuals
  ytil <- y - yhat
  dtil <- D - Dhat
  # doubly robust quantity for every sample
  drhat <- yhat1 - yhat0 + (y - yhat) * (D / Dhat - (1 - D) / (1 - Dhat))
  coef_est <- mean(drhat)
  vari <- var(drhat)
  se <- sqrt(vari / nrow(X))
  cat("point", coef_est)
  cat("se", se)
  return(list(coef_est = coef_est, se = se, ytil = ytil, dtil = dtil, drhat = drhat,
              yhat0 = yhat0, yhat1 = yhat1, dhat = Dhat, yhat = yhat))
}
```

```{r}
summaryIRM <- function(coef_est, se, ytil, dtil, drhat, name) {
  summary_data <- data.frame(
    estimate = coef_est, # point estimate
    se = se, # standard error
    lower = coef_est - 1.96 * se, # lower end of 95% confidence interval
    upper = coef_est + 1.96 * se, # upper end of 95% confidence interval
    rmse_y = sqrt(mean(ytil^2)), # res of model that predicts outcome y
    rmse_D = sqrt(mean(dtil^2)), # res of model that predicts treatment D
    accuracy_D = mean(abs(dtil) < 0.5) # binary classification accuracy of model for D
  )
  row.names(summary_data) <- name
  return(summary_data)
}
```

#### Repeat analysis in the IRM setting.

```{r}
# DML with Lasso/Logistic
set.seed(123)
cat(sprintf("\nDML with LassoCV/Logistic \n"))

dreg_lasso_cv <- function(x, d) {
  cv.glmnet(x, d, family = "binomial", alpha = 0, nfolds = 5)
}
yreg0_lasso_cv <- function(x, y) {
  cv.glmnet(x, y, family = "gaussian", alpha = 1, nfolds = 5)
}
yreg1_lasso_cv <- function(x, y) {
  cv.glmnet(x, y, family = "gaussian", alpha = 1, nfolds = 5)
}

# more folds seems to help stabilize finite sample performance
dml2_results <- dml2_for_irm(X, D, y, dreg_lasso_cv, yreg0_lasso_cv, yreg1_lasso_cv, nfold = 5)
sum_lasso_cv <- summaryIRM(dml2_results$coef_est, dml2_results$se, dml2_results$ytil, dml2_results$dtil,
                           dml2_results$drhat, name = "LassoCVLogistic")
tableirm <- data.frame()
tableirm <- rbind(sum_lasso_cv)
tableirm

yhat0_lasso <- dml2_results$yhat0
yhat1_lasso <- dml2_results$yhat1
dhat_lasso <- dml2_results$dhat
yhat_lasso <- dml2_results$yhat
```

```{r}
# DML with Random Forest
set.seed(123)
cat(sprintf("\nDML with Random Forest \n"))

dreg_rf <- function(x, d) {
  randomForest(x, d, ntree = 1000, nodesize = 10)
} # ML method=Forest
yreg0_rf <- function(x, y) {
  randomForest(x, y, ntree = 1000, nodesize = 10)
} # ML method=Forest
yreg1_rf <- function(x, y) {
  randomForest(x, y, ntree = 1000, nodesize = 10)
} # ML method=Forest


dml2_results <- dml2_for_irm(as.matrix(X), D, y, dreg_rf, yreg0_rf, yreg1_rf, nfold = 5, method = "randomforest")
sum_rf <- summaryIRM(dml2_results$coef_est, dml2_results$se, dml2_results$ytil, dml2_results$dtil,
                     dml2_results$drhat, name = "Random Forest")
tableirm <- rbind(tableirm, sum_rf)
tableirm

yhat0_rf <- dml2_results$yhat0
yhat1_rf <- dml2_results$yhat1
dhat_rf <- dml2_results$dhat
dhat_rf <- dml2_results$yhat
```

```{r}
# DML with Decision Trees
set.seed(123)
cat(sprintf("\nDML with Decision Trees \n"))

dreg_tr <- function(x, d) {
  rpart(as.formula("D~."), cbind(data.frame(D = d), x), method = "class", minbucket = 10, cp = 0.001)
}
yreg0_tr <- function(x, y) {
  rpart(as.formula("y~."), cbind(data.frame(y = y), x), minbucket = 10, cp = 0.001)
}
yreg1_tr <- function(x, y) {
  rpart(as.formula("y~."), cbind(data.frame(y = y), x), minbucket = 10, cp = 0.001)
}

dml2_results <- dml2_for_irm(X, D, y, dreg_tr, yreg0_tr, yreg1_tr, nfold = 5, method = "decisiontrees")
sum_tr <- summaryIRM(dml2_results$coef_est, dml2_results$se, dml2_results$ytil, dml2_results$dtil,
                     dml2_results$drhat, name = "Decision Trees")
tableirm <- rbind(tableirm, sum_tr)
tableirm

yhat0_tr <- dml2_results$yhat0
yhat1_tr <- dml2_results$yhat1
dhat_tr <- dml2_results$dhat
yhat_tr <- dml2_results$yhat
```

```{r}
# DML with Boosted Trees
set.seed(123)
cat(sprintf("\nDML with Boosted Trees \n"))

# NB: early stopping cannot easily be implemented with gbm
## set n.trees = best, where best <- gbm.perf(dreg_boost, plot.it = FALSE)
dreg_boost <- function(x, d) {
  gbm(as.formula("D~."), cbind(data.frame(D = d), x), distribution = "bernoulli",
      interaction.depth = 2, n.trees = 100, shrinkage = .1)
}
yreg0_boost <- function(x, y) {
  gbm(as.formula("y~."), cbind(data.frame(y = y), x), distribution = "gaussian",
      interaction.depth = 2, n.trees = 100, shrinkage = .1)
}
yreg1_boost <- function(x, y) {
  gbm(as.formula("y~."), cbind(data.frame(y = y), x), distribution = "gaussian",
      interaction.depth = 2, n.trees = 100, shrinkage = .1)
}

# passing these through regression as type="response", and D should not be factor!
dml2_results <- dml2_for_irm(X, D, y, dreg_boost, yreg0_boost, yreg1_boost, nfold = 5, method = "boostedtrees")
sum_boost <- summaryIRM(dml2_results$coef_est, dml2_results$se, dml2_results$ytil, dml2_results$dtil,
                        dml2_results$drhat, name = "Boosted Trees")
tableirm <- rbind(tableirm, sum_boost)
tableirm

yhat0_boost <- dml2_results$yhat0
yhat1_boost <- dml2_results$yhat1
dhat_boost <- dml2_results$dhat
yhat_boost <- dml2_results$yhat
```

```{r}
# Ensembles

# Best
# We'll look at model that does best for Y overall. Could also use different model for Y0 and Y1
# Here, the best performance for Y is the random forest and for D the boosted tree

# residuals
ytil <- y - dhat_rf
dtil <- D - dhat_boost
# doubly robust quantity for every sample
drhat <- yhat1_rf - yhat0_rf + (y - dhat_rf) * (D / dhat_boost - (1 - D) / (1 - dhat_boost))
coef_est <- mean(drhat)
vari <- var(drhat)
se <- sqrt(vari / nrow(X))

sum_best <- summaryIRM(coef_est, se, ytil, dtil, drhat, name = "Best")
tableirm <- rbind(tableirm, sum_best)
tableirm
```

```{r}
# Least squares model average
# We'll look at weights that do best job for Y overall. Could also use different weights for Y0 and Y1

ma_dw <- lm(D ~ dhat_lasso + dhat_rf + dhat_tr + dhat_boost)$coef
ma_yw <- lm(y ~ yhat_lasso + dhat_rf + yhat_tr + yhat_boost)$coef

Dhats <- cbind(as.matrix(rep(1, nrow(X))), dhat_lasso, dhat_rf, dhat_tr, dhat_boost)
Y0s <- cbind(as.matrix(rep(1, nrow(X))), yhat0_lasso, yhat0_rf, yhat0_tr, yhat0_boost)
Y1s <- cbind(as.matrix(rep(1, nrow(X))), yhat1_lasso, yhat1_rf, yhat1_tr, yhat1_boost)

dhat <- Dhats %*% as.matrix(ma_dw)
yhat0 <- Y0s %*% as.matrix(ma_yw)
yhat1 <- Y1s %*% as.matrix(ma_yw)

# Prediction of treatment and outcome for observed instrument
yhat <- yhat0 * (1 - D) + yhat1 * D
# residuals
ytil <- y - yhat
dtil <- D - dhat
# doubly robust quantity for every sample
drhat <- yhat1 - yhat0 + (y - yhat) * (D / dhat - (1 - D) / (1 - dhat))
coef_est <- mean(drhat)
vari <- var(drhat)
se <- sqrt(vari / nrow(X))

sum.ma <- summaryIRM(coef_est, se, ytil, dtil, drhat, name = "Model Average")
tableirm <- rbind(tableirm, sum.ma)
tableirm
```

## Double ML package

We are interested in valid estimators of the average treatment effect of `e401` and `p401` on `net_tfa`. There exist nice packages out there that can help us do our estimation with the simple call of a function. Such packages include `EconML` (Python) and `DoubleML` (Python and R).

We run through PLR and IRM using `DoubleML` below to illustrate. The `DoubleML` package internally builds on `mlr3`. We use the meta package `mlr3` to generate predictions with machine learning methods. A comprehensive introduction and description of the `mlr3` package is provided in the [mlr3book](https://mlr3book.mlr-org.com/). A list of all learners that you can use in `mlr3` can be found [here](https://mlr3extralearners.mlr-org.com/articles/learners/list_learners.html). The entry in the columns *mlr3 Package* and *Packages* indicate which packages must be installed/loaded in your R session.

You find additional information about `DoubleML` on the package on the package website https://docs.doubleml.org/ and the R documentation page https://docs.doubleml.org/r/stable/.

```{r}
install.packages("DoubleML")
install.packages("mlr3learners")
install.packages("mlr3")
install.packages("data.table")
install.packages("randomForest")
install.packages("ranger")

library(DoubleML)
library(mlr3learners)
library(mlr3)
library(data.table)
library(randomForest)
library(ranger)
```

```{r}
# Constructing the data (as DoubleMLData)
formula_flex <- paste("net_tfa ~ e401 + poly(age, 6, raw=TRUE) + poly(inc, 8, raw=TRUE) ",
                      "+ poly(educ, 4, raw=TRUE) + poly(fsize, 2, raw=TRUE) + marr + twoearn + db + pira + hown")
model_flex <- as.data.table(model.frame(formula_flex, pension))
x_cols <- colnames(model_flex)[-c(1, 2)]
data_ml <- DoubleMLData$new(model_flex, y_col = "net_tfa", d_cols = "e401", x_cols = x_cols)

p <- dim(model_flex)[2] - 2
p
```

As mentioned, in the tutorial we use the meta package `mlr3` to generate predictions with machine learning methods. A comprehensive introduction and description of the `mlr3` package is provided in the [mlr3book](https://mlr3book.mlr-org.com/). A list of all learners that you can use in `mlr3` can be found [here](https://mlr3extralearners.mlr-org.com/articles/learners/list_learners.html). The entry in the columns *mlr3 Package* and *Packages* indicate which packages must be installed/loaded in your R session.

## Partially Linear Regression Models (PLR)

```{r}
# Estimating the PLR
lgr::get_logger("mlr3")$set_threshold("warn")
lasso <- lrn("regr.cv_glmnet", nfolds = 5, s = "lambda.min")
lasso_class <- lrn("classif.cv_glmnet", nfolds = 5, s = "lambda.min")

dml_plr <- DoubleMLPLR$new(data_ml, ml_l = lasso, ml_m = lasso_class, n_folds = 5)
dml_plr$fit(store_predictions = TRUE)
dml_plr$summary()
lasso_plr <- dml_plr$coef
lasso_std_plr <- dml_plr$se
```

Let us check the predictive performance of this model.

```{r}
dml_plr$params_names()
g_hat <- as.matrix(dml_plr$predictions$ml_l) # predictions of g_o
m_hat <- as.matrix(dml_plr$predictions$ml_m) # predictions of m_o
```

```{r}
# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
theta <- as.numeric(dml_plr$coef) # estimated regression coefficient
d <- as.matrix(pension$e401)
predictions_y <- as.matrix(d * theta) + g_hat # predictions for y
lasso_y_rmse <- sqrt(mean((y - predictions_y)^2))
lasso_y_rmse
```

```{r}
# cross-fitted RMSE: treatment
d <- as.matrix(pension$e401)
lasso_d_rmse <- sqrt(mean((d - m_hat)^2))
lasso_d_rmse

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

Then, we repeat this procedure for various machine learning methods.

```{r}
# Random Forest
lgr::get_logger("mlr3")$set_threshold("warn")
randomForest <- lrn("regr.ranger")
random_forest_class <- lrn("classif.ranger")

dml_plr <- DoubleMLPLR$new(data_ml, ml_l = randomForest, ml_m = random_forest_class, n_folds = 5)
dml_plr$fit(store_predictions = TRUE) # set store_predictions=TRUE to evaluate the model
dml_plr$summary()
forest_plr <- dml_plr$coef
forest_std_plr <- dml_plr$se
```

We can compare the accuracy of this model to the model that has been estimated with lasso.

```{r}
# Evaluation predictions
g_hat <- as.matrix(dml_plr$predictions$ml_l) # predictions of g_o
m_hat <- as.matrix(dml_plr$predictions$ml_m) # predictions of m_o
theta <- as.numeric(dml_plr$coef) # estimated regression coefficient
predictions_y <- as.matrix(d * theta) + g_hat # predictions for y
forest_y_rmse <- sqrt(mean((y - predictions_y)^2))
forest_y_rmse

# cross-fitted RMSE: treatment
forest_d_rmse <- sqrt(mean((d - m_hat)^2))
forest_d_rmse

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

```{r}
# Trees
lgr::get_logger("mlr3")$set_threshold("warn")

trees <- lrn("regr.rpart")
trees_class <- lrn("classif.rpart")

dml_plr <- DoubleMLPLR$new(data_ml, ml_l = trees, ml_m = trees_class, n_folds = 5)
dml_plr$fit(store_predictions = TRUE)
dml_plr$summary()
tree_plr <- dml_plr$coef
tree_std_plr <- dml_plr$se

# Evaluation predictions
g_hat <- as.matrix(dml_plr$predictions$ml_l) # predictions of g_o
m_hat <- as.matrix(dml_plr$predictions$ml_m) # predictions of m_o
theta <- as.numeric(dml_plr$coef) # estimated regression coefficient
predictions_y <- as.matrix(d * theta) + g_hat # predictions for y
tree_y_rmse <- sqrt(mean((y - predictions_y)^2))
tree_y_rmse

# cross-fitted RMSE: treatment
tree_d_rmse <- sqrt(mean((d - m_hat)^2))
tree_d_rmse

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

```{r}
# needed to run boosting
remotes::install_github("mlr-org/mlr3extralearners")
install.packages("mlr3extralearners")
install.packages("mboost")
library(mlr3extralearners)
library(mboost)
```

```{r}
# Boosting
boost <- lrn("regr.glmboost")
boost_class <- lrn("classif.glmboost")

dml_plr <- DoubleMLPLR$new(data_ml, ml_l = boost, ml_m = boost_class, n_folds = 5)
dml_plr$fit(store_predictions = TRUE)
dml_plr$summary()
boost_plr <- dml_plr$coef
boost_std_plr <- dml_plr$se

# Evaluation predictions
g_hat <- as.matrix(dml_plr$predictions$ml_l) # predictions of g_o
m_hat <- as.matrix(dml_plr$predictions$ml_m) # predictions of m_o
theta <- as.numeric(dml_plr$coef) # estimated regression coefficient
predictions_y <- as.matrix(d * theta) + g_hat # predictions for y
boost_y_rmse <- sqrt(mean((y - predictions_y)^2))
boost_y_rmse

# cross-fitted RMSE: treatment
boost_d_rmse <- sqrt(mean((d - m_hat)^2))
boost_d_rmse

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

Let's sum up the results:

```{r}
table <- matrix(0, 4, 4)
table[1, 1:4] <- c(lasso_plr, forest_plr, tree_plr, boost_plr)
table[2, 1:4] <- c(lasso_std_plr, forest_std_plr, tree_std_plr, boost_std_plr)
table[3, 1:4] <- c(lasso_y_rmse, forest_y_rmse, tree_y_rmse, boost_y_rmse)
table[4, 1:4] <- c(lasso_d_rmse, forest_d_rmse, tree_d_rmse, boost_d_rmse)
rownames(table) <- c("Estimate", "Std.Error", "RMSE Y", "RMSE D")
colnames(table) <- c("Lasso", "Random Forest", "Trees", "Boosting")
tab <- xtable(table, digits = 2)
tab
```

The best model with lowest RMSE is the PLR model estimated via lasso (or boosting based on the RSME Y). It gives the following estimate:

```{r}
lasso_plr
```

## Interactive Regression Model (IRM)

```{r}
lgr::get_logger("mlr3")$set_threshold("warn")
dml_irm <- DoubleMLIRM$new(data_ml,
  ml_g = lasso,
  ml_m = lasso_class,
  trimming_threshold = 0.01, n_folds = 5
)
dml_irm$fit(store_predictions = TRUE)
dml_irm$summary()
lasso_irm <- dml_irm$coef
lasso_std_irm <- dml_irm$se


# predictions
dml_irm$params_names()
g0_hat <- as.matrix(dml_irm$predictions$ml_g0) # predictions of g_0(D=0, X)
g1_hat <- as.matrix(dml_irm$predictions$ml_g1) # predictions of g_0(D=1, X)
g_hat <- d * g1_hat + (1 - d) * g0_hat # predictions of g_0
m_hat <- as.matrix(dml_irm$predictions$ml_m) # predictions of m_o
```

```{r}
# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
d <- as.matrix(pension$e401)
lasso_y_irm <- sqrt(mean((y - g_hat)^2))
lasso_y_irm

# cross-fitted RMSE: treatment
lasso_d_irm <- sqrt(mean((d - m_hat)^2))
lasso_d_irm

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

```{r}
##### forest #####

dml_irm <- DoubleMLIRM$new(data_ml,
  ml_g = randomForest,
  ml_m = random_forest_class,
  trimming_threshold = 0.01, n_folds = 5
)
dml_irm$fit(store_predictions = TRUE)
dml_irm$summary()
forest_irm <- dml_irm$coef
forest_std_irm <- dml_plr$se

# predictions
g0_hat <- as.matrix(dml_irm$predictions$ml_g0) # predictions of g_0(D=0, X)
g1_hat <- as.matrix(dml_irm$predictions$ml_g1) # predictions of g_0(D=1, X)
g_hat <- d * g1_hat + (1 - d) * g0_hat # predictions of g_0
m_hat <- as.matrix(dml_irm$predictions$ml_m) # predictions of m_0

# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
d <- as.matrix(pension$e401)
forest_y_irm <- sqrt(mean((y - g_hat)^2))
forest_y_irm

# cross-fitted RMSE: treatment
forest_d_irm <- sqrt(mean((d - m_hat)^2))
forest_d_irm

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)

##### trees #####

dml_irm <- DoubleMLIRM$new(data_ml,
  ml_g = trees, ml_m = trees_class,
  trimming_threshold = 0.01, n_folds = 5
)
dml_irm$fit(store_predictions = TRUE)
dml_irm$summary()
tree_irm <- dml_irm$coef
tree_std_irm <- dml_irm$se

# predictions
g0_hat <- as.matrix(dml_irm$predictions$ml_g0) # predictions of g_0(D=0, X)
g1_hat <- as.matrix(dml_irm$predictions$ml_g1) # predictions of g_0(D=1, X)
g_hat <- d * g1_hat + (1 - d) * g0_hat # predictions of g_0
m_hat <- as.matrix(dml_irm$predictions$ml_m) # predictions of m_o

# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
d <- as.matrix(pension$e401)
tree_y_irm <- sqrt(mean((y - g_hat)^2))
tree_y_irm

# cross-fitted RMSE: treatment
tree_d_irm <- sqrt(mean((d - m_hat)^2))
tree_d_irm

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)


##### boosting #####

dml_irm <- DoubleMLIRM$new(data_ml,
  ml_g = boost, ml_m = boost_class,
  trimming_threshold = 0.01, n_folds = 5
)
dml_irm$fit(store_predictions = TRUE)
dml_irm$summary()
boost_irm <- dml_irm$coef
boost_std_irm <- dml_irm$se

# predictions
g0_hat <- as.matrix(dml_irm$predictions$ml_g0) # predictions of g_0(D=0, X)
g1_hat <- as.matrix(dml_irm$predictions$ml_g1) # predictions of g_0(D=1, X)
g_hat <- d * g1_hat + (1 - d) * g0_hat # predictions of g_0
m_hat <- as.matrix(dml_irm$predictions$ml_m) # predictions of m_o

# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
d <- as.matrix(pension$e401)
boost_y_irm <- sqrt(mean((y - g_hat)^2))
boost_y_irm

# cross-fitted RMSE: treatment
boost_d_irm <- sqrt(mean((d - m_hat)^2))
boost_d_irm

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

```{r}
table <- matrix(0, 4, 4)
table[1, 1:4] <- c(lasso_irm, forest_irm, tree_irm, boost_irm)
table[2, 1:4] <- c(lasso_std_irm, forest_std_irm, tree_std_irm, boost_std_irm)
table[3, 1:4] <- c(lasso_y_irm, forest_y_irm, tree_y_irm, boost_y_irm)
table[4, 1:4] <- c(lasso_d_irm, forest_d_irm, tree_d_irm, boost_d_irm)
rownames(table) <- c("Estimate", "Std.Error", "RMSE Y", "RMSE D")
colnames(table) <- c("Lasso", "Random Forest", "Trees", "Boosting")
tab <- xtable(table, digits = 2)
tab
```

Here, Random Forest gives the best prediction rule for $g_0$ and Lasso the best prediction rule for $m_0$, respectively. Let us fit the IRM model using the best ML method for each equation to get a final estimate for the treatment effect of eligibility.

```{r}
lgr::get_logger("mlr3")$set_threshold("warn")
dml_irm <- DoubleMLIRM$new(data_ml,
  ml_g = randomForest,
  ml_m = lasso_class,
  trimming_threshold = 0.01, n_folds = 5
)
dml_irm$fit(store_predictions = TRUE)
dml_irm$summary()
best_irm <- dml_irm$coef
best_std_irm <- dml_irm$se
```

These estimates that flexibly account for confounding are
substantially attenuated relative to the baseline estimate (*19559*) that does not account for confounding. They suggest much smaller causal effects of 401(k) eligiblity on financial asset holdings.

