---
title: An R Markdown document converted from "PM4/r-identification-analysis-of-401-k-example-w-dags.irnb"
output: html_document
---

# Using Dagitty in the Analysis of Impact of 401(k) on Net Financial Wealth

```{r}
# install and load package
install.packages("dagitty")
install.packages("ggdag")
```

```{r}
library(dagitty)
library(ggdag)
```

# Graphs for 401(K) Analsyis

Here we have
 * $Y$ -- net financial assets;
 * $X$ -- worker characteristics (income, family size, other retirement plans; see lecture notes for details);
 * $F$ -- latent (unobserved) firm characteristics
 * $D$ -- 401(K) eligibility, deterimined by $F$ and $X$

**One graph (where F determines X):**

```{r}
# generate a DAGs and plot them

G1 <- dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [uobserved, pos="0, -1"]
D -> Y
X -> D
F -> X
F -> D
X -> Y}')


ggdag(G1) + theme_dag()
```

**List minimal adjustment sets to identify causal effects $D \to Y$**


```{r}
adjustmentSets(G1, "D", "Y", effect = "total")
```

**What is the underlying principle?**

Here conditioning on X blocks backdoor paths from Y to D (Pearl).  Dagitty correctly finds X (and does many more correct decisions when we consider more elaborate structures. Why do we want to consider more elaborate structures? The very empirical problem requires us to do so!)

**Another Graph (wherere $X$ determines $F$):**

```{r}
# generate a couple of DAGs and plot them

G2 <- dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [uobserved, pos="0, -1"]
D -> Y
X -> D
X -> F
F -> D
X -> Y}')


ggdag(G2) + theme_dag()
```

```{r}
adjustmentSets(G2, "D", "Y", effect = "total")
```

**One more graph (encompassing the previous ones), where (F, X) are jointly determined by latent factors $A$.**

We can allow in fact the whole triple (D, F, X) to be jointly determined by latent factors $A$.

This is much more realistic graph to consider.

```{r}
G3 <- dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [unobserved, pos="0, -1"]
A [unobserved, pos="-1, -1"]
D -> Y
X -> D
F -> D
A -> F
A -> X
A -> D
X -> Y}')

adjustmentSets(G3, "D", "Y", effect = "total")

ggdag(G3) + theme_dag()
```

# Threat to Identification:

What if $F$ also directly affects $Y$? (Note that there are no valid adjustment sets in this case.)

```{r}
G4 <- dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [unobserved, pos="0, -1"]
A [unobserved, pos="-1, -1"]
D -> Y
X -> D
F -> D
A -> F
A -> X
A -> D
F -> Y
X -> Y}')


ggdag(G4) + theme_dag()
```

```{r}
adjustmentSets(G4, "D", "Y", effect = "total")
```

**Note that no output means that there is no valid adustment set (among observed variables).**

**How can F affect Y directly? Is it reasonable?**

Introduce Match Amount $M$. The match amount is a potential important mediator (why mediator?). $M$ is not observed. Luckily, adjusting for $X$ still works if there is no arrow $F \to M$.

```{r}
G5 <- dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [unobserved, pos="0, -1"]
A [unobserved, pos="-1, -1"]
M [unobserved, pos="2, -.5"]
D -> Y
X -> D
F -> D
A -> F
A -> X
A -> D
D -> M
M -> Y
X -> M
X -> Y}')

print(adjustmentSets(G5, "D", "Y", effect = "total"))

ggdag(G5) + theme_dag()
```

If there is an $F \to M$ arrow, then adjusting for $X$ is not sufficient.

```{r}
G6 <- dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [unobserved, pos="0, -1"]
A [unobserved, pos="-1, -1"]
M [uobserved, pos="2, -.5"]
D -> Y
X -> D
F -> D
A -> F
A -> X
D -> M
F -> M
A -> D
M -> Y
X -> M
X -> Y}')

print(adjustmentSets(G6, "D", "Y"), effect = "total")

ggdag(G6) + theme_dag()
```

Again, note that no output was returned for the adjustment set. There is no valid adjustment set here.

 # Question:

Given the analysis above, do you find the adjustment for workers' characteristics a credible strategy to identify the causal (total effect) of 401 (k) elligibility on net financial wealth?

