---
title: An R Markdown document converted from "PM4/r_dml_inference_for_gun_ownership.irnb"
output: html_document
---

# A Case Study: The Effect of Gun Ownership on Gun-Homicide Rates

We consider the problem of estimating the effect of gun ownership on the homicide rate. For this purpose, we perform inference on $\beta$ in the following the partially linear model:
$$
Y_{j, t}=\beta D_{j,(t-1)}+g\left(X_{j, t}, \bar{X}_j, \bar{X}_t, X_{j, 0}, Y_{j, 0}, t\right)+\epsilon_{j, t}
$$
$Y_{j, t}$ is the log homicide rate in county $j$ at time $t. D_{j, t-1}$ is the log fraction of suicides committed with a firearm in county $j$ at time $t-1$, which we use as a proxy for gun ownership $G_{j, t}$, which is not observed. $X_{j, t}$ is a set of demographic and economic characteristics of county $j$ at time $t$. We use $\bar{X}_j$ to denote the within county average of $X_{j, t}$ and $\bar{X}_t$ to denote the within time period average of $X_{j, t} . X_{j, 0}$ and $Y_{j, 0}$ denote initial conditions in county $j$. We use $Z_{j, t}$ to denote the set of observed control variables $\left\{X_{j, t}, \bar{X}_j, \bar{X}_t, X_{j, 0}, Y_{j, 0}, t\right\}$, so that our model is

$$
 Y_{i,t} = \beta D_{i,(t-1)} + g(Z_{i,t}) + \epsilon_{i,t}.
$$

## Data

$Y_{j,t}$ is the log homicide rate in county $j$ at time $t$, $D_{j, t-1}$ is the log fraction of suicides committed with a firearm in county $j$ at time $t-1$, which we use as a proxy for gun ownership,  and  $Z_{j,t}$ is a set of demographic and economic characteristics of county $j$ at time $t$. Assuming the firearm suicide rate is a good proxy for gun ownership, the parameter $\beta$ is the effect of gun ownership on homicide rates, controlling for county-level demographic and economic characteristics.

The sample covers 195 large United States counties between the years 1980 through 1999, giving us 3900 observations.

```{r}
install.packages("glmnet")
install.packages("randomForest")
install.packages("xgboost")
install.packages("keras")
install.packages("tensorflow")
install.packages("xtable")
install.packages("dplyr")
install.packages("sandwich")

library(glmnet)
library(randomForest)
library(xgboost)
library(keras)
library(tensorflow)
library(xtable)
library(dplyr)
library(sandwich)
```

```{r}
file <- "https://raw.githubusercontent.com/CausalAIBook/MetricsMLNotebooks/main/data/gun_clean.csv"
data <- read.csv(file)
dim(data)
```

## Preprocessing

To attempt to flexibly account for fixed heterogeneity across counties, common time factors, and deterministic time trends, we include county-level averages, time period averages, initial conditions, and the time index as additional control variables. This strategy is related to strategies for addressing latent sources of heterogeneity via conditioning.

We first reweight time and county variables as the data are population weighted.

```{r}
# Note: These data are population weighted. Specifically,
# looking at the JBES replication files, they seem to be multiplied
# by sqrt((1/T sum_t population_{j,t})/100000). To get the
# unweighted variables need to divide by this number - which we can
# get from the time effects. We are mostly just going to use the weighted
# variables as inputs - except for time and county. We'll take
# cross-sectional and time series means of these weighted variables
# as well. Note that there is nothing wrong with this, but it does not
# reproduce a weighted regression in a setting where covariates may
# enter nonlinearly and flexibly.

## County FE
county.vars <- select(data, starts_with('X_J'))

## Time variables and population weights
# Pull out time variables
time.vars <- select(data, starts_with('X_T'))

# Use these to construct population weights
popweights <- rowSums(time.vars)

# Unweighted time variables
time.vars <- time.vars/popweights

# For any columns with only zero (like the first one), just drop
time.vars <- time.vars[, colSums(time.vars != 0) > 0]

# Create time index
time.ind <- rowSums(time.vars*(seq(1:20)))
```

Now we create initial conditions, county-level averages, and time period averages.

```{r}
  ###### Create new data frame with variables we'll use

# Function to find variable names
varlist <- function (df=NULL,type=c("numeric","factor","character"), pattern="", exclude=NULL) {
  vars <- character(0)
  if (any(type %in% "numeric")) {
    vars <- c(vars,names(df)[sapply(df,is.numeric)])
  }
  if (any(type %in% "factor")) {
    vars <- c(vars,names(df)[sapply(df,is.factor)])
  }
  if (any(type %in% "character")) {
    vars <- c(vars,names(df)[sapply(df,is.character)])
  }
  vars[(!vars %in% exclude) & grepl(vars,pattern=pattern)]
}

# census control variables
census     <- NULL
census_var <- c("^AGE", "^BN", "^BP", "^BZ", "^ED", "^EL","^HI", "^HS", "^INC", "^LF", "^LN", "^PI", "^PO", "^PP", "^PV", "^SPR", "^VS")

for(i in 1:length(census_var)){
  census  <- append(census, varlist(data, pattern=census_var[i]))
}

# other control variables
X1    <- c("logrobr", "logburg", "burg_missing", "robrate_missing")
X2    <- c("newblack", "newfhh", "newmove", "newdens", "newmal")

# "treatment" variable
d     <- "logfssl"

# outcome variable
y     <- "logghomr"

# new data frame for time index
usedata <- as.data.frame(time.ind)
colnames(usedata) <- "time.ind"
usedata[,"weights"] <- popweights

varlist <- c(y,d,X1,X2,census)
for(i in 1:length(varlist)){
  usedata[, varlist[i]] <- data[,varlist[i]]
}

####################### Construct county specific means,
# time specific means, initial conditions

# Initial conditions
varlist0 <- c(y,X1,X2,census)
for(i in 1:length(varlist0)){
  usedata[, paste(varlist0[i],"0" , sep="")] <- kronecker(usedata[time.ind == 1,varlist0[i]],
                                                        rep(1,20))
}

# County means
varlistJ <- c(X1,X2,census)
county.vars <- as.matrix(county.vars)
for(i in 1:length(varlistJ)){
  usedata[, paste(varlistJ[i],"J" , sep="")] <-
    county.vars%*%qr.solve(county.vars,as.matrix(usedata[,varlistJ[i]]))
}

# Time means
time.vars <- as.matrix(time.vars)
for(i in 1:length(varlistJ)){
  usedata[, paste(varlistJ[i],"T" , sep="")] <-
    time.vars%*%qr.solve(time.vars,as.matrix(usedata[,varlistJ[i]]))
}
```

# Estimation


## Baseline OLS Estimates

After preprocessing the data, as a baseline model, we first look at simple regression of $Y_{j,t}$ on $D_{j,t-1}$ without controls in the full data set.

```{r}
# Simple regression
lm0 <- lm(logghomr ~ logfssl, data = usedata)
vc0 <- vcovHC(lm0)
cat("Baseline OLS:",lm0$coefficients[2]," (",sqrt(vc0[2,2]),")\n")
# Confidence Interval with HC3 covariance
tt <- qt(c(0.025,0.975),summary(lm0)$df[2])
se <- sqrt(diag(vc0))
ci <- coef(lm0) + se %o% tt
cat("2.5%: ", ci[2,1],"97.5%: ", ci[2,2])
```

The point estimate is $0.302$ with the confidence interval ranging from 0.277 to 0.327. This
suggests that increases in gun ownership rates are related to gun homicide rates - if gun ownership increases by 1% then the predicted gun homicide rate goes up by 0.3%, without controlling for counties' characteristics.


Next we estimate with the baseline set of controls.

```{r}
# Regression on baseline controls
varlist <- c(d,X1,X2,census)
lmC <- lm(paste("logghomr ~",paste(varlist, collapse = "+")), data = usedata)
vcC <- vcovHC(lmC)
cat("OLS with Controls:",lmC$coefficients["logfssl"]," (",sqrt(vcC["logfssl","logfssl"]),")\n")
```

<!-- Since our goal is to estimate the effect of gun ownership after controlling for a rich set county characteristics, we next include time and space averages. -->

We can also run our regression with time and space averages as controls.

```{r}
# Regression on time and cross sectional averages
varlistX <- c(X1,X2,census)
varlistMeans <- c(d,X1,X2,census)
for(i in 1:length(varlistX)){
  varlistMeans <- c(varlistMeans,paste(varlistX[i],"J" , sep=""))
}
for(i in 1:length(varlistX)){
  varlistMeans <- c(varlistMeans,paste(varlistX[i],"T" , sep=""))
}
lmM <- lm(paste("logghomr ~",paste(varlistMeans, collapse = "+")), data = usedata)
vcM <- vcovHC(lmM)
cat("OLS with Averages:",lmM$coefficients["logfssl"]," (",sqrt(vcM["logfssl","logfssl"]),")\n")
```

Since our goal is to estimate the effect of gun ownership after controlling for a rich set county characteristics, we now include all controls.

```{r}
# Regression on all controls
lmA <- lm(logghomr ~ ., data = usedata)
vcA <- vcovHC(lmA)
cat("OLS All:",lmA$coefficients["logfssl"]," (",sqrt(vcA["logfssl","logfssl"]),")\n")
```

After controlling for a rich set of characteristics, the point estimate of gun ownership attenuates to 0.179.

***NB***: In the background, `lm()` is dropping variables based on collinearity diagnostics. These depend on system linear algebra routines and can lead to large differences in high-dimensional or other ill-conditioned settings when using otherwise identical code across languages and/or machines.

Now we turn to our double machine learning framework, employing linear and flexible estimation methods with cross-fitting.

## DML Estimates

We perform inference on $\beta$ in the following the partially linear model:
 $$
Y_{j, t}=\beta D_{j,(t-1)}+g(Z_{j,t})+\epsilon_{j, t}.
$$
In the first stage, using cross-fitting, we employ modern regression methods to build estimators $\hat \ell(Z_{j,t})$ and $\hat m(Z_{j,t})$, where
- $\ell(Z_{j,t}):=E(Y_{j,t}|Z_{j,t})$
- $m(Z_{j,t}):=E(D_{j,t}|Z_{j,t})$

Using these, we obtain the estimates of the residualized quantities
- $\tilde Y_{j,t} = Y_{j,t}- E(Y_{j,t}|Z_{j,t})$
- $\tilde D_{j,t}= D_{j,t}- E(D_{j,t}|Z_{j,t})$

Using these residualized quantities, we note our model can be written as
$$
\tilde Y_{j,t} = \beta \tilde D_{j,t} + \epsilon_{j,t}, \quad E (\epsilon_{j,t} |\tilde D_{j,t}) =0.
$$
In the final stage, using ordinary least squares of $\tilde Y_{j,t}$ on $\tilde D_{j,t}$, we obtain the
estimate of $\beta$.

In the following, we consider 10 different methods for the first-stage models for $\ell(\cdot)$ and $m(\cdot)$ covering linear, penalized linear, and flexible methods. We also report the first-stage RMSE scores for estimating $Y$ and $D$.

```{r}
# NB: this cell takes > 3 hours to runon colab. To reduce computation time,
# reduce the number of cross-fitting folds. Note this may affect stability
# of estimates.

set.seed(123)

# Cross-fitting
n <- nrow(usedata)
Kf <- 5 # Number of cross-fitting folds
sampleframe <- rep(1:Kf, ceiling(n/Kf))
cvgroup <- sample(sampleframe, size=n, replace = FALSE) # Cross-fitting groups

# Initialize variables for cross-fit predictions
yhat.r <- matrix(NA,n,10)  # Going to consider 10 learners
dhat.r <- matrix(NA,n,10)

# Cross-fitting loop
for(k in 1:Kf) {
  cat("fold: ", k,"\n")
  indk <- cvgroup == k

  ktrain <- usedata[!indk,]
  ktest <- usedata[indk,]

  #### Simple regression models ####

  # Simple regression
  yhat.r[indk,1] <- ktest$logghomr - mean(ktrain$logghomr)
  dhat.r[indk,1] <- ktest$logfssl - mean(ktrain$logfssl)

  # Baseline controls
  varlist <- c(X1,X2,census)
  lmyk.C <- lm(paste("logghomr ~",paste(varlist, collapse = "+")), data = ktrain)
  yhat.r[indk,2] <- ktest$logghomr - predict(lmyk.C, ktest)
  lmdk.C <- lm(paste("logfssl ~",paste(varlist, collapse = "+")), data = ktrain)
  dhat.r[indk,2] <- ktest$logfssl - predict(lmdk.C, ktest)

  # All controls
  lmyk.A <- lm(logghomr ~ .-logfssl, data = ktrain)
  yhat.r[indk,3] <- ktest$logghomr - predict(lmyk.A, ktest)
  lmdk.A <- lm(logfssl ~ .-logghomr, data = ktrain)
  dhat.r[indk,3] <- ktest$logfssl - predict(lmdk.A, ktest)

  #### Penalized Linear Models ####

  # Lasso - default CV tuning
  ytrain <- as.matrix(usedata[!indk,"logghomr"])
  dtrain <- as.matrix(usedata[!indk,"logfssl"])
  xtrain <- as.matrix(usedata[!indk,!names(usedata) %in%
                      c("logghomr", "logfssl")])
  ytest <- as.matrix(usedata[indk,"logghomr"])
  dtest <- as.matrix(usedata[indk,"logfssl"])
  xtest <- as.matrix(usedata[indk,!names(usedata) %in%
                      c("logghomr", "logfssl")])

  lassoyk <- cv.glmnet(xtrain,ytrain)
  yhat.r[indk,4] <- ytest - predict(lassoyk, newx = xtest, s = "lambda.min")

  lassodk <- cv.glmnet(xtrain,dtrain)
  dhat.r[indk,4] <- dtest - predict(lassodk, newx = xtest, s = "lambda.min")

  # Ridge
  ridgeyk <- cv.glmnet(xtrain,ytrain,alpha = 0)
  yhat.r[indk,5] <- ytest - predict(ridgeyk, newx = xtest, s = "lambda.min")

  ridgedk <- cv.glmnet(xtrain,dtrain, alpha = 0)
  dhat.r[indk,5] <- dtest - predict(ridgedk, newx = xtest, s = "lambda.min")

  # EN, .5 - no cv over alpha
  enyk <- cv.glmnet(xtrain,ytrain,alpha = .5)
  yhat.r[indk,6] <- ytest - predict(enyk, newx = xtest, s = "lambda.min")

  endk <- cv.glmnet(xtrain,dtrain, alpha = .5)
  dhat.r[indk,6] <- dtest - predict(endk, newx = xtest, s = "lambda.min")

  #### Flexible regression models ####

  # Random forest
  rfyk <- randomForest(logghomr ~ .-logfssl, data = ktrain)
  yhat.r[indk,7] <- ktest$logghomr - predict(rfyk, ktest)
  rfdk <- randomForest(logfssl ~ .-logghomr, data = ktrain)
  dhat.r[indk,7] <- ktest$logfssl - predict(rfdk, ktest)

  # Boosted tree - depth 4
  xgb_train.y = xgb.DMatrix(data = as.matrix(usedata[!indk,!names(usedata) %in%
                                             c("logghomr", "logfssl")]),
                            label = as.matrix(usedata[!indk,"logghomr"]))
  xgb_test.y = xgb.DMatrix(data = as.matrix(usedata[indk,!names(usedata) %in%
                                            c("logghomr", "logfssl")]),
                           label = as.matrix(usedata[indk,"logghomr"]))
  xgb_train.d = xgb.DMatrix(data = as.matrix(usedata[!indk,!names(usedata) %in%
                                             c("logghomr", "logfssl")]),
                            label = as.matrix(usedata[!indk,"logfssl"]))
  xgb_test.d = xgb.DMatrix(data = as.matrix(usedata[indk,!names(usedata) %in%
                                            c("logghomr", "logfssl")]),
                           label = as.matrix(usedata[indk,"logfssl"]))

  byk = xgb.cv(data = xgb_train.y,
               nrounds = 1000, verbose = 0, eta = .1, max_depth = 4, nfold = 5)
  best.iter = which.min(as.matrix(byk$evaluation_log[,4]))
  byk = xgboost(data = xgb_train.y,
                nrounds = 1000, verbose = 0, eta = .1, max_depth = 4)
  yhat.r[indk,8] = ktest$logghomr - predict(byk, newdata = xgb_test.y,
                                            iterationrange = c(1,(best.iter+1)))

  bdk = xgb.cv(data = xgb_train.d,
               nrounds = 1000, verbose = 0, eta = .1, max_depth = 4, nfold = 5)
  best.iter = which.min(as.matrix(bdk$evaluation_log[,4]))
  bdk = xgboost(data = xgb_train.d,
                nrounds = 1000, verbose = 0, eta = .1, max_depth = 4)
  dhat.r[indk,8] = ktest$logfssl - predict(bdk, newdata = xgb_test.d,
                                            iterationrange = c(1,(best.iter+1)))

  #### Neural Networks  ####

  # normalize the covariate data
  mean <- apply(xtrain, 2, mean)
  std <- apply(xtrain, 2, sd)
  xtrainNN <- scale(xtrain, center = mean, scale = std)
  xtestNN <- scale(xtest, center = mean, scale = std)

  xtestNN <- xtestNN[,which(!is.nan(colMeans(xtrainNN)))]
  xtrainNN <- xtrainNN[,which(!is.nan(colMeans(xtrainNN)))]

  # DNN 50/50/50/50, .5 dropout
  NNmodely <- keras_model_sequential()
  NNmodely %>% layer_dense(units = 50, activation = 'relu', input_shape = c(ncol(xtrainNN))) %>%
    layer_dropout(rate = .5) %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dropout(rate = .5) %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dropout(rate = .5) %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dropout(rate = .5) %>%
    layer_dense(units = 1)

  NNmodely %>% compile(
    loss = "mse",
    optimizer = optimizer_rmsprop())

  fit.NNmodely <- NNmodely %>% fit(
    xtrainNN, ytrain,
    epochs = 200, batch_size = 200,
    validation_split = .2, verbose = 0
  )
  yhat.r[indk,9] <- ktest$logghomr - predict(NNmodely, xtestNN)

  NNmodeld <- keras_model_sequential()
  NNmodeld %>% layer_dense(units = 50, activation = 'relu', input_shape = c(ncol(xtrainNN))) %>%
    layer_dropout(rate = .5) %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dropout(rate = .5) %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dropout(rate = .5) %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dropout(rate = .5) %>%
    layer_dense(units = 1)

  NNmodeld %>% compile(
    loss = "mse",
    optimizer = optimizer_rmsprop())

  fit.NNmodeld <- NNmodeld %>% fit(
    xtrainNN, dtrain,
    epochs = 200, batch_size = 200,
    validation_split = .2, verbose = 0
  )
  dhat.r[indk,9] <- ktest$logfssl - predict(NNmodeld, xtestNN)

  # DNN 50/50/50/50, early stopping
  NNmodely <- keras_model_sequential()
  NNmodely %>% layer_dense(units = 50, activation = 'relu', input_shape = c(ncol(xtrainNN))) %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dense(units = 1)

  NNmodely %>% compile(
    loss = "mse",
    optimizer = optimizer_rmsprop())

  early.stop <- callback_early_stopping(monitor = "val_loss", patience = 25,
                                        restore_best_weights = TRUE)

  fit.NNmodely <- NNmodely %>% fit(
    xtrainNN, ytrain,
    epochs = 200, batch_size = 200,
    validation_split = .2, verbose = 0,
    callbacks = list(early.stop)
  )
  yhat.r[indk,10] <- ktest$logghomr - predict(NNmodely, xtestNN)

  NNmodeld <- keras_model_sequential()
  NNmodeld %>% layer_dense(units = 50, activation = 'relu', input_shape = c(ncol(xtrainNN))) %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dense(units = 50, activation = 'relu') %>%
    layer_dense(units = 1)

  NNmodeld %>% compile(
    loss = "mse",
    optimizer = optimizer_rmsprop())

  early.stop <- callback_early_stopping(monitor = "val_loss", patience = 25,
                                        restore_best_weights = TRUE)

  fit.NNmodeld <- NNmodeld %>% fit(
    xtrainNN, dtrain,
    epochs = 200, batch_size = 200,
    validation_split = .2, verbose = 0,
    callbacks = list(early.stop)
  )
  dhat.r[indk,10] <- ktest$logfssl - predict(NNmodeld, xtestNN)

}

################################################################################
# Predictions done, now DML

RMSE.y <- sqrt(colMeans(yhat.r^2))
RMSE.d <- sqrt(colMeans(dhat.r^2))

# dml coefficient estimates
b.dml <- rep(NA,10)
s.dml <- rep(NA,10)
for(k in 1:10){
  lm.k <- lm(yhat.r[,k] ~ dhat.r[,k]-1)
  v.k <- vcovHC(lm.k)
  b.dml[k] <- lm.k$coefficients
  s.dml[k] <- sqrt(v.k)
}

# "best" coefficient estimate
lm.k <- lm(yhat.r[,which.min(RMSE.y)] ~ dhat.r[,which.min(RMSE.d)]-1)
v.k <- vcovHC(lm.k)
b.dml[11] <- lm.k$coefficients
s.dml[11] <- sqrt(v.k)

# ls model average
yhat <- usedata$logghomr - yhat.r
dhat <- usedata$logfssl - dhat.r

ma.y <- lm(usedata$logghomr ~ yhat-1)
ma.d <- lm(usedata$logfssl ~ dhat-1)
weights.y <- ma.y$coefficients
weights.d <- ma.d$coefficients
lm.k <- lm(ma.y$residuals ~ ma.d$residuals-1)
v.k <- vcovHC(lm.k)
b.dml[12] <- lm.k$coefficients
s.dml[12] <- sqrt(v.k)

## Display results
table1 <- matrix(0, 10, 2)
table1[,1]   <- RMSE.y
table1[,2]   <- RMSE.d
colnames(table1)<- c("RMSE Y","RMSE D")
rownames(table1)<- c("OLS - No Controls", "OLS - Basic", "OLS - All",
                    "Lasso (CV)", "Ridge (CV)", "Elastic Net (.5,CV)",
                    "Random Forest","Boosted trees - depth 4",
                    "DNN - 50/50/50/50, dropout", "DNN - 50/50/50/50, early stopping")
tab1<- xtable(table1, digits =c(0,4,4))
tab1

table2 <- matrix(0, 12, 2)
table2[,1]   <- b.dml
table2[,2]   <- s.dml
colnames(table2)<- c("Point Estimate","Std. Error")
rownames(table2)<- c("OLS - No Controls", "OLS - Basic", "OLS - All",
                    "Lasso (CV)", "Ridge (CV)", "Elastic Net (.5,CV)",
                    "Random Forest","Boosted trees - depth 4",
                    "DNN - 50/50/50/50, dropout", "DNN - 50/50/50/50, early stopping",
                    "Best","Least Squares Model Average")
tab2<- xtable(table2, digits =c(0,4,4))
tab2
```

```{r}
print(xtable(table1,type="latex"))
print(xtable(table2,type="latex"))
```

